### 2023-01-05 TemporalGraphCube/old-version/TemporalGraphCube_full.tex

%% bare_jrnl_compsoc.tex

\documentclass[10pt,journal,compsoc]{IEEEtran}
%\usepackage{ntheorem}
%\usepackage{algorithmic}
%\usepackage{cite}
%\usepackage{times}
\usepackage{amsmath,amsfonts,amssymb, amsthm, xspace}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{graphicx, subfigure,stfloats}
\usepackage{epstopdf}
\usepackage{textcomp}
\usepackage{ragged2e, xcolor}
\usepackage{balance}
\usepackage{verbatim}
\usepackage{enumitem}
\usepackage{makecell}
\usepackage{url}
\usepackage{booktabs}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[10pt,journal,compsoc]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
% IEEE Computer Society needs nocompress option
% requires cite.sty v4.0 or later (November 2003)
\usepackage[nocompress]{cite}
\else
% normal IEEE
\usepackage{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later. Note also the use of a CLASSOPTION conditional provided by
% IEEEtran.cls V1.7 and later.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
% \usepackage[pdftex]{graphicx}
% declare the path(s) where your graphic files are
% \graphicspath{{../pdf/}{../jpeg/}}
% and their extensions so you won't have to specify these with
% every instance of \includegraphics
% \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
% or other class option (dvipsone, dvipdf, if not using dvips). graphicx
% will default to the driver specified in the system graphics.cfg if no
% driver is specified.
% \usepackage[dvips]{graphicx}
% declare the path(s) where your graphic files are
% \graphicspath{{../eps/}}
% and their extensions so you won't have to specify these with
% every instance of \includegraphics
% \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at:
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** SPECIALIZED LIST PACKAGES ***

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a sans serif font rather
% than the serif font used in traditional IEEE formatting and thus the need
% to invoke different subfig.sty package options depending on whether
% compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and
% Axel Sommerfeldt. This package may be useful when used in conjunction with
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\renewcommand{\baselinestretch}{0.99}

\input{command}

\begin{document}
	%
	% paper title
	% Titles are generally capitalized except for words such as a, an, and, as,
	% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
	% not capitalized unless they are the first or last word of the title.
	% Linebreaks \\ can be used within to get better formatting as desired.
	% Do not put math or special symbols in the title.
	\title{Temporal Graph Cube}
	%
	%
	% author names and IEEE memberships
	% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
	% a structure at a ~ so this keeps an author's name from being broken across
	% two lines.
	% use \thanks{} to gain access to the first footnote area
	% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
	% was not built to handle multiple paragraphs
	%
	%
	%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
	% lists the Computer Society journals use for "first footnote" author
	% affiliations. Use \IEEEcompsocthanksitem which works much like \item
	% for each affiliation group. When not in compsoc mode,
	% \IEEEcompsocitemizethanks becomes like \thanks and
	% \IEEEcompsocthanksitem becomes a line break with idention. This
	% facilitates dual compilation, although admittedly the differences in the
	% desired content of \author between the different types of papers makes a
	% one-size-fits-all approach a daunting prospect. For instance, compsoc
	% journal papers have the author affiliations above the "Manuscript
	% received ..."  text while in non-compsoc journals this is reversed. Sigh.

	\author{Qiangqiang~Dai, Rong-Hua~Li, Guoren~Wang, Rui~Mao, Zhiwei~Zhang, Ye~Yuan
		\IEEEcompsocitemizethanks{
			\IEEEcompsocthanksitem Q.~Dai, R.-H.~Li, G.~Wang, Z.~Zhang and Y.~Yuan are with the Beijing Institute of Technology, Beijing 100081, China. E-mail: qiangd66@gmail.com, lironghuascut@gmail.com, wanggrbit@126.com, cszwzhang@outlook.com, yuanye@mail.neu.edu.cn.
			\IEEEcompsocthanksitem R.~Mao is with the Shenzhen University, Shenzhen 518060, China. E-mail: mao@szu.edu.cn.
		}
	}

	%\author{Michael~Shell,~\IEEEmembership{Member,~IEEE,}
		%        John~Doe,~\IEEEmembership{Fellow,~OSA,}
		%        and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
		%\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem M. Shell was with the Department
			%of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
			%GA, 30332.\protect\\
			%% note need leading \protect in front of \\ to get a newline within \thanks as
			%% \\ is fragile and will error, could use \hfil\break instead.
			%E-mail: see http://www.michaelshell.org/contact.html
			%\IEEEcompsocthanksitem J. Doe and J. Doe are with Anonymous University.}% <-this % stops an unwanted space
		%%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}
		%}

	% note the % following the last \IEEEmembership and also \thanks -
	% these prevent an unwanted space from occurring between the last author name
	% and the end of the author line. i.e.., if you had this:
	%
	% \author{....lastname \thanks{...} \thanks{...} }
	%                     ^------------^------------^----Do not want these spaces!
	%
	% a space would be appended to the last name and could cause every name on that
	% line to be shifted left slightly. This is one of those "LaTeX things". For
	% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
	% "AB" then you have to do: "\textbf{A}\textbf{B}"
	% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers IEEE Transactions on Knowledge and Data Eengineering
\markboth{IEEE Transactions on Knowledge and Data Eengineering, ,~Vol.~X, No.~X, XXXX~2021}%
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
%
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2015 IEEE}%
	%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society jorunal
% papers don't need this extra clearance.)

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
	\begin{abstract}
		\justifying We study the problem of enabling data warehouse and OLAP technology to analyze large scale temporal multidimensional networks. Data warehouse and OLAP (Online Analytical Processing) are effective tools for decision support on traditional relational data and static multidimensional network. However, real-world networks can usually be modeled as temporal multidimensional networks and they can not be handled by data warehouse and OLAP technology yet. In this paper, we propose a new data warehouse model, \kw{Temporal\enspace Graph\enspace Cube}. By extending the basic idea of static graph cube, \kw{Temporal\enspace Graph\enspace Cube} supports OLAP queries on temporal multidimensional networks. Specifically, by supporting OLAP queries in any time range, users can obtain summative information of the networks in the time range of interest, which is impossible with static graph cube. In conducting OLAP queries of \kw{Temporal\enspace Graph\enspace Cube}, we should merge snapshots in a certain time range. We abstract this problem as range query problem on snapshot array, and we extend three typical indexes in commong range query problem: prefix array, sparse table and segment tree to accelerate the query. We enable these indexes to be updated in real time, because the size of real-world temporal networks grow over time. We find that the above indexes are not always work, and sometimes the baseline method is the best method. Based on this, we define \kw{similarity\enspace of\enspace snapshots}, which is proved to have a strong correlation with the efficiency of indexes and can provide a reference on the necessity to build indexes. In implementing \kw{Temporal\enspace Graph\enspace Cube}, we find that the traditional algorithms used to implement traditional data cube and static graph cube are no longer applicable to \kw{Temporal\enspace Graph\enspace Cube}. Instead, we adopt another simple method \kw{MinLevel}, which is introduced in implementation of static graph cube. The results of extensive experiments on real-world and man-made datasets show the effectiveness of \kw{Temporal\enspace Graph\enspace Cube} and corroborate the correlation between \kw{similarity\enspace of\enspace snapshots} and the efficiency of indexes. With the support of \kw{MinLevel} strategy, the indexes are able to achieve good acceleration effect and space efficiency, especially segment tree, which also has the best scalability.
	\end{abstract}

	% Note that keywords are not normally used for peerreview papers.
	\begin{IEEEkeywords}
		Uncertain Graphs, Cohesive Subgraph Mining, Uncertain Core Decomposition.
\end{IEEEkeywords}}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEtitleabstractindextext text will
% not be used in maketitle, but will appear (i.e.., to be "transported")
% here as \IEEEdisplaynontitleabstractindextext when the compsoc
% or transmag modes are not selected <OR> if conference mode is selected
% - because all conference papers position the abstract like regular
% papers do.
\IEEEdisplaynontitleabstractindextext
% \IEEEdisplaynontitleabstractindextext has no effect when using
% compsoc or transmag under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}

Data warehouses and OLAP (Online Analytical Processing) technology have proven to be helpful tools for knowledge workers (executive, manager, analyst) to analyze and make decisions, because Data warehouses and OLAP together can efficiently provide summarized information in different resolutions by specifying different views (different combinations of data dimensions) of large-scale real-world data through OLAP operations such as roll-up, drill-down and slice-and-dice \cite{chaudhuri1997overview}. Data warehouse built on traditional RDB data can be called data cube \cite{gray1997data}. Zhao et al. \cite{zhao2011graph} proposed \kw{Graph} \kw{Cube} in 2011 which extends data warehouses and OLAP techniques so that they can be used to analyze static multidimensional networks. For each query, static graph cube returns a static network, and the summative information is in it's structure and statistical values on vertices and edges, unlike traditional data cube which only returns the statistical values obtained by group-by operations. However, many real-life networks, such as human proximity networks, scientific collaboration networks, and biological networks can be modeled as temporal networks \cite{holme2015modern}, where each relationship or interaction has a corresponding timestamp. Also, vertices in temporal networks usually contain attributes of multiple dimensions, for example, vertices in human proximity networks have attributes such as name, gender, nationality, hobbies, etc. Therefore, we can further construct the above networks as temporal multidimensional networks. There has been many works focusing on the management and analysis of temporal networks \cite{junghanns2017management}, but none of them have tried to extend OLAP techniques to temporal multidimensional networks, i.e., developing ways to provide users in real time with summative information about temporal multidimensional networks in different resolutions and time ranges.
\begin{figure*}[t!]
	\begin{center}
	\begin{tabular}[t]{c}
		\centering
		\subfigure[{\scriptsize Vertex table}]{
			\includegraphics[width=0.27\textwidth]{fig/example/example_tables/tb_example_dataset_adjust.pdf}
		}
		\subfigure[{\scriptsize Network structure}]{
			\includegraphics[width=0.63\textwidth]{fig/example/example_dataset_trade.pdf}
		}
	\end{tabular}
	\end{center}
	\vspace*{-0.4cm}
	\caption{A sample temporal multidimensional network with its vertex table and network structure}
	\vspace*{-0.4cm}
	\label{fig:example_dataset}
\end{figure*}

\begin{example} \label{exam:temp_multi_network}
	Figure~\ref{fig:example_dataset} shows a sample temporal transaction network, presenting the transactions between individuals of different countries of birth and professions in each day. Table in Figure~\ref{fig:example_dataset} (a) is the vertex table of the temporal network, showing the information of the 8 individuals. Each individual has a primary key ID and 3 dimensions (3 discrete attributes): gender, country (country of birth), profession. Income is a numeric attribute of individuals. Figure~\ref{fig:example_dataset} (b) shows the network structrue of the temporal network. There are 20 temporal edges, illustrating the transactions between individuals in 5 days. Temporal edges in each day can be organized into a snapshot. The numeric attribute on each temporal edge is the amount of each transaction. The static multidimensional information of individuals in Figure~\ref{fig:example_dataset} (a) and the temporal information of network structure in Figure~\ref{fig:example_dataset} (b) form a temporal multidimensional network.
\end{example}
In static graph cube, users can specify different views to obtain summative information about the static multidimensional network and combine the information under these views through OLAP operations such as roll-up, drill-down and slice-and-dice interactively for decision support and business intelligence. For example, a company tries to use static graph cube to analyze the interaction characteristics of people with different attributes (name, gender, profession, income level, etc.) in a large-scale social network in order to support their marketing strategy making. However, edges in real-world networks usually contain timestamps. For example, in social networks, interactions between people occur at specific times, such as a particular day. In this way, static graph cube does not work when users only need information for a specific time range. Taking the previous example, the company only needs data from the last few years, because data that is too old is less informative. Or, they want to study what the characteristics of people's interactions are during particular time ranges, such as holidays or COVID-19 periods. Or, they want to compare several different time ranges to see how people's interaction patterns change over time. Static graph cube can not meet the above needs. As a result, we need to extend static graph cube to make data warehouses and OLAP capable of analyzing temporal multidimensional networks.

\begin{figure}[t!]
	\begin{center}
		\begin{tabular}[t]{c}
			\centering
			\hspace*{+0.3cm}
			\subfigure[{\scriptsize Vertex table}]{
				\includegraphics[width=0.36\columnwidth]{fig/example/example_tables/tb_cuboid_gender_adjust.pdf}
			}
			\hspace*{-0.6cm}
			\subfigure[{\scriptsize Network structure}]{
				\includegraphics[width=0.58\columnwidth]{fig/example/exam_cuboid_gender_2_3.pdf}
			}
			\hspace*{-0.6cm}
		\end{tabular}
	\end{center}
	\vspace*{-0.4cm}
	\caption{The answer to ``What is the amount of transactions between individuals of different genders from day 2 to day 3?"}
	\vspace*{-0.4cm}
	\label{fig:exam_cuboid_gender}
\end{figure}
\begin{figure}[t!]
	\begin{center}
		\includegraphics[width=0.45\columnwidth]{fig/example/exam_cuboid_gender_semi_state.pdf}
	\end{center}
	\vspace*{-0.7cm}
	\caption{Intermediate result of generating network structure in Figure~\ref{fig:exam_cuboid_gender} (b)}
	\vspace*{-0.4cm}
	\label{fig:exam_inter_result_gender}
\end{figure}

%\begin{table}[t!]%\vspace*{-0.3cm}
%	%\scriptsize
%	\small
%	\centering
%	\caption{Vertex table of the first question in Example~\ref{exam:cuboid_query}} \vspace*{-0.2cm} \label{tab:cuboid_gender}
%	\begin{tabular}{c c}
%		\toprule
%		Gender & COUNT($\ast$) \cr %\hline %\hline
%		\midrule
%		male&4 \cr %\hline
%		female&4  \cr %\hline
%		\bottomrule
%	\end{tabular}
%	\vspace*{0.4cm}
%\end{table}

%\begin{figure*}[t!]
%	\begin{center}
%		\includegraphics[width=0.95\textwidth]{fig/example/exam_aggregate_gender.pdf}
%	\end{center}
%	\vspace*{-0.4cm}
%	\caption{example aggregate gender}
%	\vspace*{-0.4cm}
%	\label{fig:aggregation_gender}
%\end{figure*}

%\begin{table}[t!]%\vspace*{-0.3cm}
%	%\scriptsize
%	\small
%	\centering
%	\caption{Vertex table of the second and the third question in Example~\ref{exam:cuboid_query}} \vspace*{-0.2cm} \label{tab:cuboid_gender_profession}
%	\begin{tabular}{c c c}
%		\toprule
%		Gender & Profession & AVERAGE($\ast$) \cr %\hline %\hline
%		\midrule
%		male & teacher& 5,000 \cr %\hline
%		male & shopowner & 10,000  \cr %\hline
%		male & student & 1,000  \cr %\hline
%		female & shopowner & 6,000  \cr %\hline
%		female & lawyer & 7,000  \cr %\hline
%		female & doctor & 9,000  \cr %\hline
%		\bottomrule
%	\end{tabular}
%	\vspace*{-0.4cm}
%\end{table}

%\begin{figure*}[t!]
%	\begin{center}
%		\includegraphics[width=0.95\textwidth]{fig/example/exam_aggregate_gender_pro.pdf}
%	\end{center}
%	\vspace*{-0.4cm}
%	\caption{example aggregate gender profession}
%	\vspace*{-0.4cm}
%	\label{fig:aggregation_gender_profession}
%\end{figure*}
%

%\begin{figure}[t!]
%	\begin{center}
%		\begin{tabular}[t]{c}\hspace*{-0.3cm}
%			\subfigure[{\scriptsize Structrue of the answer of the first question in Example~\ref{exam:cuboid_query}}]{
%				\includegraphics[width=0.45\columnwidth]{fig/example/exam_cuboid_gender_2_3_adjust.pdf}
%			}
%			\subfigure[{\scriptsize The intermediate result of generating (a)}]{
%				\includegraphics[width=0.45\columnwidth]{fig/example/exam_cuboid_gender_semi_state.pdf}
%			}
%		\end{tabular}
%	\end{center}
%	\vspace*{-0.4cm}
%	\caption{}
%	\vspace*{-0.4cm}
%	\label{fig:exam_cuboid_gender}
%\end{figure}

\begin{figure}[t!]
	\begin{center}
		\begin{tabular}[t]{c}\hspace*{-0.2cm}
			\subfigure[{\scriptsize Vertex table}]{
				\includegraphics[width=0.4\columnwidth]{fig/example/example_tables/tb_cuboid_gender_pro_adjust.pdf}
			}
			\subfigure[{\scriptsize Network structure from day 2 to day 3}]{
				\includegraphics[width=0.5\columnwidth]{fig/example/exam_cuboid_gender_pro_2_3.pdf}
			}
			%\vspace*{-0.3cm}\\\hspace*{-0.3cm}
			\\
			\subfigure[{\scriptsize Network structure from day 4 to day 5}]{
				\includegraphics[width=0.5\columnwidth]{fig/example/exam_cuboid_gender_pro_4_5.pdf}
			}
		\end{tabular}
	\end{center}
	\vspace*{-0.4cm}
	\caption{The answers to ``What is the amount of transactions between individuals of different genders and professions from day 2 to day 3 and from day 4 to day 5?"}
	\vspace*{-0.4cm}
	\label{fig:exam_cuboid_gender_profession}
\end{figure}
\begin{example} \label{exam:cuboid_query}
	Figure~\ref{fig:exam_cuboid_gender} shows a static aggregate network obtained by summarizing the temporal multidimensional network in Figure~\ref{fig:example_dataset} on dimension ``Gender'' and time range $ [2,3] $. Figure~\ref{fig:exam_cuboid_gender} can be regarded as the answer of question ``What is the amount of transactions between individuals of different genders from day 2 to day 3?" Note that the vertex table in Figure~\ref{fig:exam_cuboid_gender} (a) is directly aggregated from the vertex table in Figure~\ref{fig:example_dataset} (a) on ``Gender" with COUNT($\ast$) as aggregation function on vertices. All vertices in Figure~\ref{fig:example_dataset} (a) are included in the aggregation process even those vertices which have no adjacent edge in day 2 and day 3, like $ v_5 $. The reason is that vertices are time independent and vertices having no adjacent edge in specified time ranges are still available in those time ranges. So from day 2 to day 3, the COUNT($\ast$) value of male should be 4. Figure~\ref{fig:exam_inter_result_gender} shows the intermediate result of the question, which is obtained by merging all snapshots in day 2 and day 3. All edges of snapshots in day 2 and day 3 are collected to the single snapshot in Figure~\ref{fig:exam_inter_result_gender}, and if there are two edges in snpshots of day 2 and day 3 respectively share the same vertices, then their numeric attributes will be added together since SUM($\ast$) is the aggregation function upon edges here. Figure~\ref{fig:exam_cuboid_gender} (b) can be directly aggregated from Figure~\ref{fig:exam_inter_result_gender} on the basis of aggregated vertices in Figure~\ref{fig:exam_cuboid_gender} (a), i.e., each edge in Figure~\ref{fig:exam_cuboid_gender} (b) is the aggregated edge of edges in Figure~\ref{fig:exam_inter_result_gender}, whose two vertices are aggregated to the two vertices of the corresponding aggregated edge. The numeric attributes of the aggregated edges in Figure~\ref{fig:exam_cuboid_gender} (b) are still the sum of attributes of their corresponding sets of edges in Figure~\ref{fig:exam_inter_result_gender}.

	Finally we can reply the question above, for example, from day 2 to day 3, the amount of transactions between males is 6, while the amount of transactions between male and female is 107, etc. Figure~\ref{fig:exam_cuboid_gender_profession} are the answers of two similar questions ``What is the amount of transactions between individuals of different genders and professions from day 2 to day 3 and from day 4 to day 5?" The only difference between these two questions is the time range specified. Both static aggregate networks in Figure~\ref{fig:exam_cuboid_gender_profession} share the same vertex table in Figure~\ref{fig:exam_cuboid_gender_profession} (a) and in these two questions the aggregation functions upon vertices are AVERAGE($\ast$).
\end{example}
There are three new OLAP queries in Example~\ref{exam:cuboid_query} compared to queries in static graph cube, for example, the first OLAP query in Eample~\ref{exam:cuboid_query} is ``What is the amount of transactions between individuals of different genders from day 2 to day 3", which requests a rough summative result of the temporal multidimensional network in Figure~\ref{fig:example_dataset} on ``Gender". Unlike the query in static graph cube, this new query includes a specific time range $ [2,3] $. Conducting the new query includes the same aggregation on vertex table as in static graph cube, but when aggregating the network structure, the basic idea is to first merge snapshots in the time range to build an intermediate result, as in Figure~\ref{fig:exam_inter_result_gender}, then perform the same aggregation on Figure~\ref{fig:exam_inter_result_gender} as in static graph cube. If the new query specifies the whole time range of the temporal multidimensional network, then it's equivalent to a query in static graph cube. Therefore, static graph cube is a special version of the new data warehouse model to be proposed in this paper. Conducting the second OLAP query ``What is the amount of transactions between individuals of different genders and professions from day 2 to day 3" can be regarded as performing a drill-down operation on the basis of the first query, since the second query requests more fine-grained information in the same time range than the first query. The third query ``What is the amount of transactions between individuals of different genders and professions from day 4 to day 5" specifies the same dimensions as the second query, but specifies a different time range $ [4,5] $. This allows users to analyze the pattern characteristics in different time ranges, for example, there is no transaction between male students from day 2 to day 3, but there are transactions totaling 100 between them from day 4 to day 5. Note that the aggregation functions upon vertices and edges can be combined arbitrarily as needed. For example, we can specify AVERAGE($\ast$) as the aggregation function upon both vertices and edges in Example~\ref{exam:cuboid_query}, so that we can analyze the transaction network from another perspective.

In \cite{zhao2011graph}, Zhao et al. proposed a new OLAP query especially for static multidimensional networks \kw{crossboid}, to analyze the crossed interaction pattern of two different views. \kw{Crossboid} can also be easily extended to the temporal multidimensional network scenario as in Example~\ref{exam:crossboid_query}.

\begin{table}[t!]%\vspace*{-0.3cm}
	%\scriptsize
	\small
	\centering
	\caption{Vertex table obtained by group-by on ``Profession" in Figure~\ref{fig:example_dataset} (a)} \vspace*{-0.2cm} \label{tab:groupby_profession}
	\begin{tabular}{c c}
		\toprule
		Profession & COUNT($\ast$) \cr %\hline %\hline
		\midrule
		teacher& 1 \cr %\hline
		shopowner & 2  \cr %\hline
		student & 2  \cr %\hline
		lawyer & 2  \cr %\hline
		doctor & 1  \cr %\hline
		\bottomrule
	\end{tabular}
	\vspace*{-0.4cm}
\end{table}

\begin{figure}[t!]
	\begin{center}
		\includegraphics[width=0.85\columnwidth]{fig/example/exam_crossboid_gender_pro_2_3.pdf}
	\end{center}
	\vspace*{-0.7cm}
	\caption{Structure of the answer to the question in Example~\ref{exam:crossboid_query}}
	\vspace*{-0.4cm}
	\label{fig:exam_crossboid_gender_profession}
\end{figure}

\begin{example} \label{exam:crossboid_query}
	Figure~\ref{fig:exam_crossboid_gender_profession} shows the structure of the answer to ``What is the amount of transactions between individuals grouped by gender and individuals grouped by profession from day 2 to day 3?" The answer can also be regarded as a static aggregate network which has two vertex tables from group-by on ``Gender" and ``Profession" shown in Figure~\ref{fig:exam_cuboid_gender} (a) and Table~\ref{tab:groupby_profession} respectively. Figure~\ref{fig:exam_crossboid_gender_profession} shows the crossed transactions between two different views ``Gender" and ``Profession" from day 2 to day 3, e.g., for all transactions between males and individuals of different professions, all 4 males have the biggest amount of transactions with the 2 laywers and the smallest amount of transactions with the 2 students from day 2 to day 3. Other crossed transactions can also be easily found in Figure~\ref{fig:exam_crossboid_gender_profession}.
\end{example}
In this paper, we propose a new data warehouse model, \kw{Temporal} \kw{Graph} \kw{Cube}, and we extend the definition of OLAP queries by specifying a time range so that they can be applied to \kw{Temporal} \kw{Graph} \kw{Cube}. Users can explore summarized information on different views of temporal multidimensional networks in any time range using \kw{Temporal} \kw{Graph} \kw{Cube}. Here, we face a new problem: how to efficiently merge snapshots in a specific time range for each online OLAP query. In Example~\ref{exam:cuboid_query}, we merge snapshots in time ranges $ [2,3] $ and $ [4,5] $ one by one. Such a basic idea is inefficient when the time ranges expand and a large number of queries come. The above problem can be seen as range query on snapshot arrays. Unfortunately, among the existing works for summarizing temporal networks \cite{liu2018graph}, there is no work which focuses on summarizing snapshots within a certain time range online. In this paper, we investigate the problem of how to speed up merging snapshots in certain time ranges, i.e., building indexes on snapshot arrays to reduce response time of queries. Specifically, we extend typical indexes that are useful in traditional range query problems: prefix array, sparse table, and segment tree to the range query problem on snapshot arrays. In addition, since new edges are constantly added to the real-world temporal multidimensional networks, we also need to consider the maintenance of \kw{Temporal} \kw{Graph} \kw{Cube} and the indexes built in it when adding new edges. As with static graph cube, the implementation of the \kw{Temporal} \kw{Graph} \kw{Cube} requires determining the materialization strategy of views in order to achieve a balance between time and space. We refer to the strategy \kw{MinLevel} proposed in \cite{zhao2011graph}, as it fits our model best.

The main contributions of our work are summarized as follows:
\begin{enumerate}
	\item We propose a new data warehouse model \kw{Temporal} \kw{Graph} \kw{Cube}, which supports decision making on the basis of temporal multidimensional networks. The biggest difference compared to static graph cube is that \kw{Temporal} \kw{Graph} \kw{Cube} supports querying summative information for any time range of temporal multidimensional networks so that it supports more diverse analysis.

	\item We extend three typical indexes that work on common range query problems to range query problems on snapshot arrays to reduce response time of queries. We also make these indexes support self-updating when new edges are added to the temporal networks. Similar requirement has not been seen on common range query problems.

	\item We find that these indexes do not always work as in common range query problem, and their effectiveness is related to some nature of the snapshot arrays. We propose a metric: \kw{similarity\enspace of\enspace snapshots}, to measure the overall similarity of snapshots in a snapshot array, or, the degree to which the edges are shared by different snapshots. We find that this metric has a strong correlation with the effectiveness of the indexes and it indicates which snapshot arrays indexes should be built on.

	\item We point out that the common methods used to implement traditional data cube and static graph cube: greedy algorithm \cite{harinarayan1996implementing} and its variants \cite{morfonios2007rolap}, are no longer applicable in \kw{Temporal \enspace Graph \enspace Cube} scenario. The main reason is that OLAP queries contain arbitrary time ranges. A more reasonable method is \kw{MinLevel}, which is very simple to be implemented and does not take into account the time ranges in OLAP queries.

	\item We conduct extensive experiments on real-world and man-made datasets. The results demonstrate the effectiveness of the \kw{Temporal \enspace Graph \enspace Cube}. The results also demonstrate the correlation between the efficiency (time and space) of the three indexes and \kw{similarity\enspace of\enspace snapshots}. We also examined the update performance of the three indexes. Among them, we find that segment tree is with the best time and space balance and scalability among the three indexes.
\end{enumerate}

%\vspace*{-0.2cm}
\stitle{Organizations.} Section~\ref{sec:temp_graph_cube} introduces the details about \kw{Temporal \enspace Graph \enspace Cube}. We extend all three indexes to snapshot array in Section~\ref{sec:indexes}. In Section~\ref{sec:effec_of_indexes} we discuss a metric, \kw{similarity\enspace of\enspace snapshots}, which affects the efficiency of indexes and how it affects. Section~\ref{sec:partial_materialization} introduces the implementation of \kw{Temporal \enspace Graph \enspace Cube}. We conduct experiments in Section~\ref{sec:experiment} and review the related works in Section~\ref{sec:related_work}. We conclude our work in Section~\ref{sec:conclusion}.

\section{Temporal Graph Cube} \label{sec:temp_graph_cube}
%\subsection{Notations and Basic Concepts} \label{subsec:notation}
%\vspace*{-0.1cm}
\subsection{Definitions}\label{subsec:definitions}
\begin{definition}[Temporal Multidimensional Network]
	\label{def:temp_multi_network}
	Let $\Gr=(\Vr, \Er, A, W)$ be an undirected temporal multidimensional network where $\Vr$ and $\Er$ are the set of nodes and edges respectively. Edges in $\Er$ are in the form of $(u, v, t, a)$, where $u$, $v$ are nodes in $\Vr$ and $t, a$ are timestamp and numeric attribute respectively attached to edges. $A=\{A_1, A_2, ... , A_n\}$ represents the dimensions of the network, or $n$ discrete attributes of nodes in $ \Vr $, i.e., $A(u)=\{A_1(u), A_2(u), ... , A_n(u)\}$ where $A_i(u)$ is the actual value of $u$ on the i-th attribute. $ W(u) $ is the numeric value of $u$.
\end{definition}
In convenience, we abbreviate temporal multidimensional network and static multidimensional network to temp-multi-network and static-multi-network respectively. Since we only consider undirected edges in this paper, if not specified, we assign $u \leq v$ in $(u, v, t, a)$ and in all following definitions.

As in Figure~\ref{fig:example_dataset}, $ A=\{Gender,Country,Profession\} $ and $ W=Income $. We don't consider ID to be an attribute of individuals because ID is only the identification of individuals. $ \Vr $ is the set of IDs of individuals in Figure~\ref{fig:example_dataset} (a) and $ \Er $ is the set of temporal edges in Figure~\ref{fig:example_dataset} (b).

A temp-multi-network can also be represented by a sequence of snapshots. 
\begin{definition}[Snapshot]
	\label{def:snapshot}
	Let $\Tr=\{t|(u, v, t, a)\in \Er\}$ be the set of timestamps. For each $t_i\in \Tr$, we can obtain a snapshot $S_i=\{(u,v,a)|(u,v,t_i,a)\in \Er\}$.
\end{definition}
For example, in Figure~\ref{fig:example_dataset} (b), we can extract 5 snapshots and snapshot at timestamp 1 $ S_1=\{(v_1,v_2,10),(v_2,v_3,20),(v_3,v_4,10),(v_1,v_4,15)\} $. With the definition of snapshot, we can define temp-multi-network in Definition~\ref{def:temp_multi_network} as $\Gr=(\Vr,\Sr,A,W)$ where $\Sr=\{S_1,S_2,...,S_{|\Tr|}\}$ and $ S_i $ is the snapshot at $t_i$.

\stitle{Biased Timestamp.} To store snapshots in a snapshot array and access snapshots directly by corresponding timestamps, we replace the original timestamps of snapshots with biased timestamps. For each temp-multi-network, suppose $ t_1 $ is the smallest timestamp, $ t_{|\Tr|} $ is the biggest timestamp and $ t_{base}=t_1-1 $, biased timestamp of snapshot $ S_i $ is computed by $ t_i-t_{base} $. Through biased timestamp we can map timestamps in $ [t_1,t_{|\Tr|}] $ to $ [1,t_{end}] $, where $ t_{end}=t_{|\Tr|}-t_{base} $. The above strategy is useful when timestamps in temp-multi-network are dense. If timestamps are sparse, i.e., $ |\Tr| \ll t_{|\Tr|}-t_1 $, there will be many empty snapshots in snapshot array. We can sacrifice a little in efficiency, hashing all the timestamps into a denser space and relocating time range $ [l,r] $ of each query (we will discuss queries in Section~\ref{subsec:queries}) to the actual time range. However, in large scale real-world temp-multi-networks, temporal relations can be built at almost all time, so the timestamps are unlikely to be sparse, which is proven in our datasets.

All algorithms in this paper are designed based on definition of temp-multi-network where $\Gr=(\Vr,\Sr,A,W)$ and $ \Sr $ is a snapshot array where biased timestamps of valid snapshots start from 1.

Aggregation, or traditional group-by, is a key operation in data warehouses and OLAP for traditional RDB data and static-multi-network. In traditional RDB data, for an aggregation in form of $(A_1',A_2',...,A_n')$ where some (or all) dimension $A_i'$ could be $\ast$ (ALL), the result is another table in which tuples in source table having the same value in all specified dimensions (NOT $\ast$) are grouped in the same tuple. In the above process, numeric attributes of tuples in source table can be summed by aggregation functions, e.g., COUNT($\ast$), AVERAGE($\ast$), MAX($\ast$) and MIN($\ast$). In static-multi-network, aggregations are not only conducting group-by on the vertex table but constructing static aggregate networks and computing the aggregation result of attributes of edges using specified aggregation function. The results of aggregations on static-multi-network, static aggregate networks, are exactly the answers of OLAP queries on static-multi-network \cite{zhao2011graph}. However, situations are different in temp-multi-network. Next, we first define temporal aggregate network, the results of aggregations on temp-multi-network and show the difference between static aggregate network and temporal aggregate network, then we describe the benefit of temporal aggregate network in conducting OLAP queries on temp-multi-network.
\begin{definition}[Temporal Aggregate Network]
	\label{def:temp_aggregate_network}
	Given a temporal multidimensional network $\Gr=(\Vr,\Sr,A,W)$ and an aggregation $A'=(A'_1,A'_2,...,A'_n)$ where $A'_i$ equals $A_i$ or $\ast$, the obtained temporal aggregate network is another temporal multidimensional network $\Gr'=(\Vr',\Sr',B,W_{\Gr'})$ where $B=\{A'_i|A'_i \neq \ast\}$ and
	\begin{enumerate}
		\item $\Vr' \subseteq \Vr$ and $\forall [v]$, an equivalence class in $\Vr$, where $ v \in \Vr $ and $[v]=\{u|B(u)=B(v),u \in \Vr\}$, $\exists v' \in \Vr'$ satisfying $v' \in [v]$ and $\nexists u' \in \Vr', u' \neq v'$ satisfying $u' \in [v]$. The weight of $v'$, $W_{\Gr'}(v')$, is the aggregation result of $W(v)$ for $v \in [v']$ obtained by aggregation function upon vertices specified by user.
%		\item For any $S'_i$ in $\Sr'$ and each $(u',v',a')$ in $S'_i$, there exists a nonempty maximum edge set $E=\{(u,v,a)|(u,v,a) \in S_i, u \in [u'] \wedge v \in [v']\; or\; u \in [v'] \wedge v \in [u']\}$, $a'$ is the aggregation result of numeric attributes ($a$) of edges in $E$ obtained by aggregation function specified by user.
		\item $\forall u',v' \in \Vr'$ where $u' \leq v'$ and any $\Sr[i], i \in [1,t_{end}]$, if there exists a nonempty maximum edge set $E=\{(u,v,a)|(u,v,a) \in \Sr[i],(u \in [u'] \wedge v \in [v']) \vee (u \in [v'] \wedge v \in [u']) \}$, then $ \exists (u',v',a') \in \Sr'[i] $, where $a'$ is the aggregation result of numeric attributes of edges in $E$ obtained by aggregation function upon edges specified by user.
	\end{enumerate}
\end{definition}
In short, conducting aggregations on temp-multi-network contains two stages: aggregating, or group-by on the vertex table as 1) in Definition~\ref{def:temp_aggregate_network} and aggregating each snapshots in the temp-multi-network as 2) in Definition~\ref{def:temp_aggregate_network}.

\begin{figure}[t!]
	\begin{center}
		\includegraphics[width=0.95\columnwidth]{fig/example/exam_aggregate_gender.pdf}
	\end{center}
	\vspace*{-0.4cm}
	\caption{Snapshots of the first temporal aggregate network in Example~\ref{exam:temporal_aggregate_network}}
	\vspace*{-0.4cm}
	\label{fig:aggregation_gender}
\end{figure}

\begin{figure}[t!]
	\begin{center}
		\includegraphics[width=0.95\columnwidth]{fig/example/exam_aggregate_gender_pro.pdf}
	\end{center}
	\vspace*{-0.5cm}
	\caption{Snapshots of the second temporal aggregate network in Example~\ref{exam:temporal_aggregate_network}. $ v_1= $ ``male, teacher", $ v_2= $ ``male, shopowner", $ v_3= $ ``male, student", $ v_4= $ ``female, shopowner", $ v_5= $ ``female, lawyer", $ v_6= $ ``female, doctor"}
	\vspace*{-0.4cm}
	\label{fig:aggregation_gender_profession}
\end{figure}

\begin{example} \label{exam:temporal_aggregate_network}
	Figure~\ref{fig:aggregation_gender} shows the snapshots of a temporal aggregate network, which is obtained by aggregating temp-multi-network in Figure~\ref{fig:example_dataset} on dimension ``Gender". The obtained temporal aggregate network is also a temp-multi-network, with vertex table shown in Figure~\ref{fig:exam_cuboid_gender} (a). There are still 5 snapshots in Figure~\ref{fig:aggregation_gender}, each of them is corresponding to the snapshot with the same timestamp in Figure~\ref{fig:example_dataset} (b). For each snapshot in Figure~\ref{fig:aggregation_gender}, each edge $ e' $ is the aggregated edge of a set of edges $ E $ in the corresponding snapshot. Each edge in $ E $ have two vertices aggregated to the two vertices of $ e' $, as 2) in Definition~\ref{def:temp_aggregate_network}. In this example, the attribute in $ e' $ is the amount of the attributes of edges in $ E $, i.e., the total transactions between individuals of two genders in each day. We can also choose other aggregation functions for edges like COUNT($\ast$), MAX($\ast$), etc.

	Figure~\ref{fig:aggregation_gender_profession} and Table in Figure~\ref{fig:exam_cuboid_gender_profession} (a) together show another temporal aggregate network by aggregating temp-multi-network in Figure~\ref{fig:example_dataset} on dimension ``Gender" and ``Profession".
\end{example}

%As in Example~\ref{exam:temporal_aggregate_network}, Figure~\ref{fig:aggregation_gender} and Figure~\ref{fig:aggregation_gender_profession} are temporal aggregate networks $ \Gr'_1 $ and $ \Gr'_2 $ aggregated from Figure~\ref{fig:example_dataset} on $ (Gender,\ast,\ast) $ and $ (Gender,\ast,Profession) $ respectively. For $ \Gr'_1=(\Vr'_1,\Sr'_1,B_1,W'_1) $, $ \Vr'_1 $ is the set of nodes, which are represented by tuples in "Gender" column of Table~\ref{tab:aggregation_gender}. $ \Sr'_1 $ is the snapshot array starting from subscript 1 and storing the 5 snapshots in Figure~\ref{fig:aggregation_gender}. $ B_1=\{Gender\} $. $ W'_1 $ is the mapping from $ \Vr'_1 $ to COUNT($\ast$) value of nodes in $ \Vr'_1 $ as shown in "COUNT($\ast$)" column in Table~\ref{tab:aggregation_gender}. The aggregation function upon edges is SUM($\ast$). For $ \Gr'_2 $, [mention it in introduction].

From Definition~\ref{def:temp_aggregate_network} and Example~\ref{exam:temporal_aggregate_network} we can find that aggregation on static-multi-network is a special case of aggregation on temp-multi-network, if there is only one snapshot in temp-multi-network. For most cases, compared to aggregation on static-multi-network, aggregation on temp-multi-network contains aggregating each snapshot of the temp-multi-network to condensed snapshot independently.

Obviously, temporal aggregate network is not the direct answer of OLAP query on temp-multi-network, because temporal aggregate network contains aggregated snapshot at each timestamp, but OLAP query on temp-multi-network requires a summative snapshot of a specified time range. However, temporal aggregate network can provide a cheaper way to conduct queries compared to conducting queries directly on the original temp-multi-network. For example, the snapshot in Figure~\ref{fig:exam_cuboid_gender} (b) can be obtained by merging snapshots in day 2 and day 3 in Figure~\ref{fig:aggregation_gender}, without generating the intermediate result in Figure~\ref{fig:exam_inter_result_gender}. We will explain in detail in Section~\ref{subsec:queries}.

\begin{algorithm}[t]
	\scriptsize
	\caption{$\kw{Temporal Aggregate Network Construction}$}
	\label{alg:view_construct_baseline}
	\KwIn{A temp-multi-network $\Gr=(\Vr,\Sr,A,W)$; Aggregation $A'$; Aggregation function $f_v$ and $f_e$ upon vertices and edges respectively}
	\KwOut{Temporal aggregate network $\Gr'=(\Vr',\Sr',B,W_{\Gr'})$, $B=\{A'_i|A'_i \neq \ast\}$}
	$ \Vr' \gets \{\} $, let $ h $ be hash structure: $ \{B(u)|u \in \Vr\} \rightarrow \Vr' $;

	\ForEach{$u \in \Vr$}{
		\If{$h(B(u))=NULL$}{
			$ \Vr' \gets \Vr' \cup \{u\} $;
			$ W_{\Gr'}(u) \gets W(u) $;
			$ h(B(u)) \leftarrow u $;
		}
		\Else{
			$ u' \gets h(B(u)) $;
			$ W_{\Gr'}(u') \gets f_v(W_{\Gr'}(u'),W(u)) $;
		}
	}
	$ \Sr' \gets $ a snapshot array with $ t_{end}+1 $ empty snapshots;

	\ForEach{$ i \in [1,t_{end}] $}{
		$ EP \gets \{\} $;

		\ForEach{$ (u,v,a) \in \Sr[i] $}{
			$u' \gets h(B(u))$;
			$v' \gets h(B(v))$;

			\lIf{$ u'>v' $}{
				$ u',v' \gets v',u' $
			}

			\If{$(u',v') \notin EP$}{
				$ EP \gets EP \cup \{(u',v')\}$;
				$ \Sr'[i] \gets \Sr'[i] \cup \{(u',v',a)\} $;
			}
			\Else{
				Let $ (u',v',a') $ be edge in $\Sr'[i]$ which has $ u',v' $ as vertices\;
				$ \Sr'[i] \gets \Sr'[i]-\{(u',v',a')\} $\;
				$ \Sr'[i] \gets \Sr'[i] \cup \{(u',v',f_e(a',a))\} $;
			}
		}
	}
	\Return $ \Gr'=(\Vr',\Sr',B,W_{\Gr'}) $;
\end{algorithm}
Algorithm~\ref{alg:view_construct_baseline} is a basic algorithm to construct a temporal aggregate network upon aggregation $ A' $ from the original network $ \Gr $. There are two things to be concerned about aggregation functions:
\begin{itemize}[leftmargin=*]
	\item We assume $ f_v $ and $ f_e $ are not AVERAGE($\ast$), since the result of AVERAGE($\ast$) can be easily computed by SUM($\ast$)/COUNT($\ast$).

	\item if $ f_v =$ COUNT($\ast$), we should assign $1$ to all $ W(u),u \in \Vr $, because $ 1 $ is the correct attribute value of each vertex when counting is needed. The same assignment should be done to numeric attributes of all edges if $ f_e= $ COUNT($\ast$).
\end{itemize}
Algorithm~\ref{alg:view_construct_baseline} first conducts the specified aggregation on all vertices of the original network. In line~1, we first create a hash structure, $h$, to maintain a mapping from all possible $ B(u),u \in \Vr $ to vertices of the resultant temporal aggregate network. In line~2-6, we group all vertices in $ \Vr $ with the specified aggregation $ A' $ and compute the aggregation result of numeric attributes of vertices with the specified aggregation function $ f_v $.

Each snapshot $ \Sr[i] $ of $ \Gr $ should be aggregated into a smaller representation, i.e., the corresponding snapshot $ \Sr'[i] $ at the same timestamp in $ \Gr' $ (line~8-18). For each edge $ (u,v,a) $, we first map the two vertices $ u,v $ to $ u',v' $ using $ h $ (line~11-12). If there exists an aggregated edge $ (u',v',a') $ in $ \Sr'[i] $, we update $ a' $ with $ a $ using $ f_e $ (line~16-18), otherwise we insert $ (u',v',a) $ into $ \Sr'[i] $ directly (line~14).

The time complexity of Algorithm~\ref{alg:view_construct_baseline} is $ O(|\Vr|+\sum_{i=1}^{t_{end}}{|\Sr[i]|}) $, which contains time used for traversing $ \Vr $ and $ \Sr $. The space used to maintain $ h,W_{\Gr'} $ is $ O(\Vr') $ and we need $ O(|\Vr'|+\sum_{i=1}^{t_{end}}{|\Sr'[i]|}) $ space to maintain $ \Gr' $. From Algorithm~\ref{alg:view_construct_baseline} we can easily find $ |\Vr'| \leq |\Vr| $ and $ |\Sr'[i]| \leq |\Sr[i]| $, so the space complexity of Algorithm~\ref{alg:view_construct_baseline} is also $ O(|\Vr|+\sum_{i=1}^{t_{end}}{|\Sr[i]|}) $.

Similar to static aggregate network in static graph cube, temporal aggregate networks can also be construct from other existing temporal aggregate networks, e.g., temporal aggregate network with dimensions $ \{B_1,B_2\} $ can be constructed from temporal aggregate network with dimensions $ \{B_1,B_2,B_3\} $, $ \{B_1,B_2,B_4\} $, etc., instead of being constructed from the large-scale original network. The proof can be extended easily from \cite{zhao2011graph}.
\begin{definition}[Temporal Graph Cube]
	\label{def:temp_graph_cube}
	Given a temporal multidimensional network $\Gr=(\Vr,\Sr,A,W)$, the temporal graph cube is obtained by decomposing $ A $ into all possible aggregations. Each aggregation $ A' $ is a node in the temporal graph cube and corresponds to a temporal aggregate network $ \Gr' $ as difined in Definition~\ref{def:temp_aggregate_network}.
\end{definition}

In \cite{zhao2011graph}, Zhao et al. use equivalently the terms \textit{cuboid}, \textit{view} and \textit{aggregation}, however in this paper, we only use \textit{view} and \textit{aggregation} equivalently, while \textit{temporal cuboid} is used to refer to temporal cuboid query. In following sections, computing a view $ A' $ means materializing the corresponding temporal aggregate network $ \Gr' $ in memory. If $ A' $ is precomputed, then for simplicity, $ A' $ also refers to the corresponding temporal aggregate network $ \Gr' $ directly.

For a view $ A' $ in temporal graph cube, similar to static cuboid in static graph cube, $ dim(A') $ denotes the set of non-$\ast$ dimensions of $ A' $, i.e., $ dim(A')=B $ where $ B $ is the discrete attributes of vertices in $ \Gr' $, the corresponding temporal aggregate network of $ A' $. For two views $ A' $ and $ A'' $, $ A' $ is an \textit{ancestor} of $ A'' $ and $ A'' $ is a \textit{descendant} of $ A' $ if $ dim(A') \subset dim(A'') $, denoted as $ A' \preceq A'' $. Especially, the base view $ A_{base} $ is a view where $ dim(A_{base})=A $, $ A $ is the dimensions of the original temp-multi-network temporal graph cube built on. $ A_{base} $ don't need to be computed and it's a descendant of all other views. Another special view is $ A_{apex}=(\ast,\ast,...,\ast) $, and it's an ancestor of all other views in the temporal graph cube.
\begin{figure}[t]
	\begin{center}
		\includegraphics[width=0.95\columnwidth]{fig/example/cube_lattice.pdf}
	\end{center}
	\vspace*{-0.5cm}
	\caption{A sample cube lattice}
	\vspace*{-0.4cm}
	\label{fig:cube_lattice}
\end{figure}
\begin{example} \label{exam:cube_lattice}
	Figure~\ref{fig:cube_lattice} shows the temporal graph cube lattice built on temp-multi-network in Figure~\ref{fig:example_dataset}, each node on the lattice represents a view. For each edge in the lattice, the upper view is an ancestor of the lower view.
\end{example}
Temporal graph cube has the same form with static graph cube. The only difference is that each node in temporal graph cube is corresponding to a temporal aggregate network, not a static aggregate network. For each temporal aggregate network corresponding to a view, it contains more coarse-grained information than that of temporal aggregate network corresponding to a descendent view at each timestamp. In temporal graph cube, users can obtain summative information in different resolution of the original temp-multi-network in certain period by traversing the temporal graph cube lattice with specified time range. Users can also stay at some nodes in the lattice and query the summative information in different time ranges. In these ways, summative information of temp-multi-network in different resolutions and time ranges can be analyzed for decision support and business intelligence purposes.

\subsection{OLAP Queries}\label{subsec:queries}
Cuboid and crossboid are two important OLAP queries on static-multi-network, or in static graph cube \cite{zhao2011graph}. In this section, we give the formal definition of OLAP queries in temporal graph cube, by extending cuboid and crossboid to temporal cuboid and temporal crossboid respectively. From the previous sections we can find that the main difference of queries in static graph cube and temporal graph cube is the designation of time range.

\stitle{Temporal cuboid.} The inputs of temporal cuboid query $ Q=(A',[l,r]) $ in temporal graph cube contain a specified view $A'=(A'_1,A'_2,...,A'_n)$ and a time range $[l,r]$. The output is a static aggregate network. As in Example~\ref{exam:cuboid_query}, the first question can be represented by $ Q=((Gender),[2,3]) $, and the static aggregate network as the answer is shown in Figure~\ref{fig:exam_cuboid_gender}.

\begin{algorithm}[t]
	\scriptsize
	\caption{$\kw{Temporal Cuboid Query}$}
	\label{alg:temp_cuboid_baseline}
	\KwIn{A temp-multi-network $\Gr=(\Vr,\Sr,A,W)$; Aggregation $A'$; Time range $[l,r]$; Aggregation function $f_v$ and $f_e$ upon vertices and edges respectively}
	\KwOut{A static aggregate network $G=(V,E,B,W_{G})$, $B=\{A'_i|A'_i \neq \ast\}$}
	$V \leftarrow \{\}$, let $h$ be hash structure: $\{B(u)|u \in \Vr\} \rightarrow V$;

	\ForEach{$u \in \Vr$}{
		\If{$h(B(u))=NULL$}{
			$ V \leftarrow V \cup \{u\} $;
			$ W_{G}(u) \gets W(u) $;
			$ h(B(u)) \leftarrow u $;
		}
		\Else{
			$ u'\gets h(B(u)) $;
			$ W_{G}(u') \gets f_v(W_{G}(u'),W(u)) $;
		}
	}
	$E' \leftarrow \{\}$;
	$EP \leftarrow \{\}$;
	$ l \gets l-t_{base} $;
	$ r \gets r-t_{base} $;

	\ForEach{$i \in [l,r]$}{
		\ForEach{$(u,v,a) \in \Sr[i]$}{
			\If{$(u,v) \notin EP$}{
				$EP \leftarrow EP \cup \{(u,v)\}$;
				$E' \leftarrow E' \cup \{(u,v,a))\}$;
			}
			\Else{
				Let $(u,v,a')$ be edge in $E'$ which has $u,v$ as vertices\;
				$E' \leftarrow E'-\{(u,v,a')\}$\;
				$E' \leftarrow E' \cup \{(u,v,f_e(a',a))\}$;
			}
		}
	}
	$EP \leftarrow \{\}$;
	$ E \gets \{\} $;

	\ForEach{$(u,v,a) \in E'$}{
		$u' \leftarrow h(B(u))$;
		$v' \leftarrow h(B(v))$;

		\lIf{$ u'>v' $}{
			$ u',v' \gets v',u' $
		}

		\If{$(u',v') \notin EP$}{
			$ EP \gets EP \cup \{(u',v')\}$;
			$ E \gets E \cup \{(u',v',a)\} $;
		}
		\Else{
			Let $ (u',v',a') $ be edge in $E$ which has $ u',v' $ as vertices\;
			$ E \gets E-\{(u',v',a')\} $\;
			$ E \gets E \cup \{(u',v',f_e(a',a))\} $;
		}
	}
	\Return $ G=(V,E,B,W_{G}) $;
\end{algorithm}
Algorithm~\ref{alg:temp_cuboid_baseline} is a basic algorithm to conduct temporal cuboid query on original temp-multi-network (or on $ A_{base} $). Same as Algorithm~\ref{alg:view_construct_baseline}, Algorithm~\ref{alg:temp_cuboid_baseline} first conducts the specified aggregation on the whole vertex table of the original network, even those vertices who don't exist in any edge of snapshots in $[l,r]$ (line~1-6). As we have mentioned in Example~\ref{exam:cuboid_query}, it's necessary to aggregate all vertices in temporal cuboid query, since vertices are time independent and they are available in all time ranges.

Process in line~1-6 is the same as line~1-6 in Algorithm~\ref{alg:view_construct_baseline}. In line~7-15, we merge all snapshots in $[l,r]$ by collecting all edges of those snapshots and summing the numeric attributes of edges sharing the same vertices using $ f_e $. Then in line~17-25, we map the two vertices of all edges in $ E' $ using $ h $ and compute the aggregation result of edges sharing the same mapped vertices using $ f_e $, which is similar to what Algorithm~\ref{alg:view_construct_baseline} do in each snapshot $ \Sr[i] $ of $ \Gr $ (line~8-18 of Algorithm~\ref{alg:view_construct_baseline}).

Before giving time and space complexity of Algorithm~\ref{alg:temp_cuboid_baseline}, we first give a brief definition of summative snapshot.
\begin{definition}[Summative Snapshot]
	\label{def:sum_snap}
	Given a temporal multidimensional network $\Gr=(\Vr,\Sr,A,W)$, the summative snapshot $ S $ of $ \Gr $ is a special snapshot of $ \Gr $, which is the network structure in the result of a temporal cuboid query $ Q=(A,[1+t_{base},t_{end}+t_{base}]) $ on $ \Gr $.
\end{definition}
In other words, summative snapshot $ S $ of $ \Gr $ can be obtained by merging all snapshots in snapshot array $ \Sr $.

The time complexity of Algorithm~\ref{alg:temp_cuboid_baseline} contains three parts: conducting aggregation on all vertices (line~1-6), merging all snapshots in specified time range into $ E' $ (line~7-15) and aggregating on $ E' $ (line~17-25). So the time complexity of Algorithm~\ref{alg:temp_cuboid_baseline} is $ O(|\Vr|+\sum_{i=l-t_{base}}^{r-t_{base}}|\Sr[i]|+|E'|) $. With the definition of summative snapshot, we can rewrite the time complexity as $ O(|\Vr|+\sum_{i=l-t_{base}}^{r-t_{base}}|\Sr[i]|+|S|) $, since we can easily find $ |E'| \leq |S|$, where $ S $ is the summative snapshot of $ \Gr $. The memory consumed by $V$, $h$ and $W_{G}$ is linearly related to $ |V| $ and $ |V| \leq |\Vr| $. The size of $ E $, $ E' $ and $ EP $ are all smaller than $ |S| $, so the space complexity of Algorithm~\ref{alg:temp_cuboid_baseline} is $ O(|\Vr|+|S|) $.

Given a temporal cuboid query $ Q=(A',[l,r]) $, if we have precomputed the view $ A' $ with the same aggregation functions, we can get the result of $ Q $ from $ \Gr' $ by merging snapshots of $ \Gr' $ in $ [l,r] $ instead.

\begin{theorem} \label{thm:cuboid_by_temp_aggre_network}
	A temporal cuboid query $ Q=(A',[l,r]) $ can be directly answered from $ A' $ if $ A' $ is precomputed with the same aggregation functions.
\end{theorem}
Theorem~\ref{thm:cuboid_by_temp_aggre_network} gives another way to conduct $ Q $. The static aggregate network as the result of $ Q $ contains two parts: vertex table and a snapshot presenting the network structure. When $ A' $ is precomputed, the vertex table of the resultant static aggregate network is the same with vertex table of $ \Gr' $, the corresponding temporal aggregate network of $ A' $, and the network structure can be obtained by merging all snapshots of $ \Gr' $ in $ [l,r] $.
\begin{proof}
	Suppose the result of $ Q=(A',[l,r]) $ is $ G=(V,E,B,W_{G}) $, and the original temp-multi-network is $ \Gr=(\Vr,\Sr,A,W) $, the temporal aggregate network of $ A' $ is $ \Gr'=(\Vr',\Sr',B,W_{\Gr'}) $. $ G $ is obtainted from $ \Gr $ by Algorithm~\ref{alg:temp_cuboid_baseline}. Since aggregating all vertices is the first step of both constructing $ \Gr' $ and obtaining $ G $, we have $ V=\Vr' $ and $ W_{G}=W_{\Gr'} $, i.e., the vertex table of $ G $ can be obtained from vertex table of $ \Gr' $ directly. As for network structure of $ G $, we only need to prove that the merging result of snapshots of $ \Gr' $ in $ [l,r] $, $ E' $, is identical to $ E $.

	Suppose $ l,r $ have been converted to biased timestamps. Recall the hash structure $ h $ used to maintain the mapping from $ \{B(u)|u \in \Vr\} $ to aggregated vertices set $ \Vr' $ and $ V $ in Algorithm~\ref{alg:view_construct_baseline} and Algorithm~\ref{alg:temp_cuboid_baseline} respectively, since $ \Vr \rightarrow \{B(u)|u \in \Vr\} $ can also be regarded as a mapping, here we can regard $ H $ as a mapping from $ \Vr $ to both $ \Vr' $ and $ V $, i.e., $ H(u)=h(B(u)) $. We first ignore the numeric attribute of each edge and prove $ E'=E $. Since $ H $ can map the two vertices of edges in $ \Sr $ to their aggregated vertices, let $ H^e $ be a mapping from edges in $ \Sr $ to their aggregated edges in $ \Sr' $ and in $ E $, i.e., $ H^e(e)=e' $ where $ e $ is an arbitrary edge in $ \Sr $ and $ e' $ is the aggregated edge of $ e $ in $ \Sr' $ or in $ E $. For $ E $, through Algorithm~\ref{alg:temp_cuboid_baseline} we know that all snapshots of $ \Sr $ in $ [l,r] $ are merged into an intermediate result $ X=\bigcup_{i=l}^{r}\Sr[i] $ first. Then $ E=\{H^e(e)|e \in X\} $. For $ E' $, we know that $ E' $ is obtained by merging all snapshots of $ \Sr' $ in $ [l,r] $ directly, i.e., $ E'=\bigcup_{i=l}^{r}\Sr'[i] $. Through Algorithm~\ref{alg:view_construct_baseline} we have $\forall i, \Sr'[i]=\{H^e(e)|e \in \Sr[i]\} $, so $ E'=\bigcup_{i=l}^{r}\{H^e(e)|e\in\Sr[i]\}=\{H^e(e)|e\in\bigcup_{i=l}^{r}\Sr[i]\} $. As a result, $ E=E' $.

	Then we take numeric attributes of edges into account and suppose the aggregation functions upon edges in $ \Gr' $ and in $ Q $ are both $ f_e $. The original form of edges in $ \Sr $ and in $ \Sr' $ is $ (u,v,a) $, suppose we add the timestamp into the form of edges as $ (u,v,a,t) $ to distinguish edges in different snapshots but with the same vertices. Each aggregated edge $ e\in E $ corresponds to a set of edges $TX_{\Sr}(e)=\{(u,v,a,i)|(u,v,a,i)\in \Sr[i],i\in [l,r],H^e((u,v))=(e.u,e.v)\} $, because $ e $ is obtained by merging and aggregating edges in $ TX_{\Sr}(e) $, and $ e.a $ is computed by executing $ f_e $ on multiset $ \{a|(u,v,a,t)\in TX_{\Sr}(e)\} $. According to the previous proof, there must exist $ e' \in E'$ where $(e'.u,e'.v)=(e.u,e.v) $. $ e' $ is obtained by merging edges in $TX_{\Sr'}(e')= \{(u,v,a,i)|(u,v,a,i) \in \Sr'[i],i \in [l,r],(u,v)=(e'.u,e'.v)\} $, and $ e'.a $ is computed by executing $ f_e $ on multiset $ \{a|(u,v,a,t)\in TX_{\Sr'}(e')\} $. Further more, each $ e'_t\in TX_{\Sr'}(e') $ where $ e'_t=(u,v,a,t) $ corresponds to a set of edges in $ \Sr[t] $, which is $TX_{\Sr[t]}(e'_t)= \{(u,v,a,t)|(u,v,a,t)\in \Sr[t],H^e((u,v))=(e'_t.u,e'_t.v)\} $. $ e'_t.a $ is computed by executing $ f_e $ on multiset $ \{a|(u,v,a,t)\in TX_{\Sr[t]}(e'_t)\} $. As a result, $ e'.a $ can be computed by executing $ f_e $ on multiset $ \{a|(u,v,a,t)\in \bigcup_{e'_t\in TX_{\Sr'}(e')}TX_{\Sr[t]}(e'_t)\}=\{a|(u,v,a,t)\in TX_{\Sr}(e')\} $. Since $ (e'.u,e'.v)=(e.u,e.v) $, we have $ TX_{\Sr}(e')=TX_{\Sr}(e) $, so $ e'.a=e.a $. Finally we prove that $ E' $, obtained by directly merging snapshots of $ \Gr' $, is equal to $ E $, the network structure in the result of $ Q $ by conducting $ Q $ on $ \Gr $.
\end{proof}

The algorithm of conducting $ Q $ on $ \Gr' $ is exactly the subprocess in line~7-15 of Algorithm~\ref{alg:temp_cuboid_baseline} if we replace $ \Gr $ with $ \Gr' $. The time complexity of the simpler algorithm is $ O(\sum_{i=l-t_{base}}^{r-t_{base}}|\Sr'[i]|) $. It's clear that $ |\Sr'[i]| \leq |\Sr[i]| $, so the new algorithm is better than the baseline algorithm.

If view $ A' $ is not precomputed while some other view $ A'' $ is precomputed with the same aggregation functions and $ dim(A') \subset dim(A'') $, we can use the corresponding temporal aggregate network $ \Gr'' $ of $ A'' $ to get the result of $ Q $. The algorithm is exactly Algorithm~\ref{alg:temp_cuboid_baseline}, where we only need replace $ \Gr $ with $ \Gr'' $

\begin{theorem} \label{thm:cuboid_by_small_temp_aggre_network}
	A temporal cuboid query $ Q=(A',[l,r]) $ can be answered from $ A'' $ where $ dim(A') \subset dim(A'') $ and $ A'' $ is precomputed with the same aggregation functions.
\end{theorem}
\begin{proof}
	Suppose $ G=(V,E,B,W_{G}) $ is the result of conducting $ Q $ from the original temp-multi-network $ \Gr=(\Vr,\Sr,A,W) $, $ \Gr''=(\Vr'',\Sr'',B'',W_{\Gr''}) $ is the temporal aggregate network corresponds to $ A'' $, $ G''=(V'',E'',B,W_{G''}) $ is the result of conducting $ Q $ from $ \Gr'' $. Three hash structures will be involved in this proof, $ h_1 $: $ \{B(u)|u \in \Vr\} \rightarrow V $, used in Algorithm~\ref{alg:temp_cuboid_baseline} when answering $ Q $ from $ \Gr $; $ h_2 $: $ \{B''(u)|u\in \Vr\} \rightarrow \Vr'' $, used in Algorithm~\ref{alg:view_construct_baseline} when constructing $ \Gr'' $ from $ \Gr $; $ h_3 $: $ \{B(u)|u\in \Vr''\} \rightarrow V'' $, used in Algorithm~\ref{alg:temp_cuboid_baseline} when answering $ Q $ from $ \Gr'' $. The key is to prove $ G=G'' $.

	We first prove that $ G $ and $ G'' $ share the same vertex table, i.e., $ T=(V,B,W_{G})$ and $T''=(V'',B,W_{G''}) $ are the same vertex table. $T$ is obtained by the traditional group-by operation on $ (\Vr,A,W) $ with aggregation $ A' $. $ T'' $ is obtained by two group-by operations: group-by on $ (\Vr,A,W) $ with $ A'' $, obtaining $ (\Vr'',B'',W_{\Gr''}) $, and group-by on $ (\Vr'',B'',W_{\Gr''}) $ with $ A' $. It's easy to know that the two group-by operations to obtain $ T'' $ is equivalent to the single group-by operation to obtain $ T $, because $ dim(A') \subset dim(A'') $. So $ T $ and $ T'' $ are the same vertex table. However there is no guarantee that $ V=V'' $, i.e., the label sets of vertices in $ T $ and $ T'' $ are not necessarily the same. The reason is that in line~2-6 of Algorithm~\ref{alg:view_construct_baseline} and Algorithm~\ref{alg:temp_cuboid_baseline}, we simply map $ B(u) $ to the first $ u $ that satisfies $ h(B(u))=NULL $ and the order $ u $ is traversed is uncertain. It's not important for group-by operations, but to prove $ E=E'' $, we need to assume that $ u $ is traversed in ascending order first ($ V=V'' $ in this case).

	Suppose the above condition of traversing $ u $ in ascending order holds, and equivalence class $ [u]_{B,\Vr}=\{v|B(u)=B(v),v\in \Vr\} $. Similar to proof of Theorem~\ref{thm:cuboid_by_temp_aggre_network}, we let $ H_1 $ be a mapping where $ H_1(u)=h_1(B(u)),u\in \Vr $ and $ H_2 $, $ H_3 $ are similar. We have $ V=\{H_1(u)|u\in \Vr\} $, $ \Vr''=\{H_2(u)|u\in \Vr\} $ and $ V''=\{H_3(u)|u\in \Vr''\} $. Since $ u $ is traversed in ascending order, we can rewrite $ H_1(u)=MIN([u]_{B,\Vr}),u\in \Vr $, $ H_2(u)=MIN([u]_{B'',\Vr}),u\in \Vr $ and $ H_3(u)=MIN([u]_{B,\Vr''}),u\in \Vr'' $, where $ MIN(\ast) $ returns the smallest label in the equivalence class. We first prove $ \forall u\in \Vr, H_1(u)=H_3(H_2(u)) $. It's clear that $ [u]_{B'',\Vr}\subseteq [u]_{B,\Vr} $ because $ B\subset B'' $ ($ dim(A')\subset dim(A'') $). Since $ H_1(u)=MIN([H_1(u)]_{B,\Vr}) $, $ [H_1(u)]_{B'',\Vr}\subseteq [H_1(u)]_{B,\Vr} $ and $ H_1(u)\in [H_1(u)]_{B'',\Vr}$, we have $ H_1(u)=MIN([H_1(u)]_{B'',\Vr}) $, so $ H_2(H_1(u))=H_1(u) $. Since $ H_2(H_1(u))=H_1(u) $, $ H_1(u) $ is in $ \Vr'' $, the vertex set of $ \Gr'' $. Since $ [u]_{B'',\Vr}\subseteq [u]_{B,\Vr} $, $ H_2(u)\in [u]_{B,\Vr} $. As a result, both $ H_1(u) $ and $ H_2(u) $ are elements in $ [u]_{B,\Vr} $ and in $ \Vr'' $, so $ B(H_1(u))=B(H_2(u)) $ and $ H_1(u)\in [H_2(u)]_{B,\Vr''} $. Since $ H_1(u) $ is the smallest label in $ \Vr $ satisfying $ B(H_1(u))=B(u) $ and $ \Vr''\subseteq \Vr $, $ H_1(u)=MIN([H_2(u)]_{B,\Vr''}) $. So $ H_3(H_2(u))=H_1(u) $.

	As we have mentioned before, $ V=\{H_1(u)|u\in \Vr\} $, $ \Vr''=\{H_2(u)|u\in \Vr\} $ and $ V''=\{H_3(u)|u\in \Vr''\} $, we can further get $ V''=\{H_3(H_2(u))|u\in \Vr\}=\{H_1(u)|u\in \Vr\} $. So $ V=V'' $. Similar to proof of Theorem~\ref{thm:cuboid_by_temp_aggre_network}, we can also define mappings between edges $ H_1^e $, $ H_2^e $ and $ H_3^e $ on the basis of $ H_1 $, $ H_2 $ and $ H_3 $. Since $ \forall u\in \Vr, H_1(u)=H_3(H_2(u)) $, we can also have $ \forall e \in \Er, H_1^e(e)=H_3^e(H_2^e(e)) $, where $ \Er $ is the edge set of $ \Gr $. Using similar process in the proof of Theorem~\ref{thm:cuboid_by_temp_aggre_network}, we can prove that $ E=E'' $. We omit the details here. In fact, the order in which $ u $ is traversed in line~2 of Algorithm~\ref{alg:view_construct_baseline} and Algorithm~\ref{alg:temp_cuboid_baseline} is not important, and $ E $ is always isomorphic to $ E'' $ because the labels of vertices have no actual meaning.
\end{proof}

The time complexity of conducting $ Q $ on $ \Gr'' $ becomes $ O(|\Vr''|+\sum_{i=l-t_{base}}^{r-t_{base}}|\Sr''[i]|+|S''|) $, where $ |\Vr''|,|\Sr''[i]| $ and $ |S''| $ are vertex set, snapshot at $ i $ and summative snapshot of $ \Gr'' $ respectively. We can also easily have $ |\Vr''| \leq |\Vr|,|\Sr''[i]| \leq |\Sr[i]|$ and $|S''| \leq |S| $ and in practice, $ |\Vr''|,|\Sr''[i]| $ and $ |S''| $ are much smaller than $ |\Vr|,|\Sr[i]| $ and $ |S| $ respectively.

If we have a set of precomputed views $ \{A_1'',A_2'',...\} $ with the same aggregation functions and for each $ A_i'' $, $ dim(A') \subset dim(A_i'') $, which one should we choose in conducting $ Q $ using Algorithm~\ref{alg:temp_cuboid_baseline}? There is a similar problem in static graph cube, in which we only need to choose the view in $ \{A_1'',A_2'',...\} $ with the smallest size, since time complexity of conducting a query $ Q=(A') $ on the basis of $ A_i'' $ in static graph cube equals the size of $ A_i'' $. It's easy to obtain the size of precomputed views of static graph cube in $ O(1) $ time.

However, in temporal graph cube, extra time range should be specified in each query, so the time complexity of conducting a query $ Q=(A',[l,r]) $ on $ A_i'' $ is not the size of $ A_i'' $. It's hard to compute the accurate time complexity of Algorithm~\ref{alg:temp_cuboid_baseline} in $ O(1) $ time. The reasons are as follows:
\begin{itemize}[leftmargin=*]
	\item If we want to compute the accumulation part ($ \sum_{i=l-t_{base}}^{r-t_{base}}|\Sr[i]| $) in the time complexity of Algorithm~\ref{alg:temp_cuboid_baseline} in $ O(1) $ time, an extra prefix array should be maintained.
	\item The size of summative snapshot $ |S| $ is just an upper bound of the number of edges visited in $ E' $ of Algorithm~\ref{alg:temp_cuboid_baseline}. To get the accurate value of $ |E'| $ as fast as possible, we may have to build some indexes to get $ E' $ first. However, it is also difficult to get $ |E'| $ in $ O(1) $ time using indexes (we will talk about building indexes in Section~\ref{sec:indexes}).
\end{itemize}
We can avoid the problem of choosing an ideal $ A_i'' $ by the partial materialization strategy \kw{MinLevel} introduced in \cite{zhao2011graph}, which we will discuss in Section~\ref{sec:partial_materialization}.
%
%Let's recall the time complexity of Algorithm~\ref{alg:temp_cuboid_baseline}: $ O(|\Vr|+\sum_{i=l-t_{base}}^{r-t_{base}}|\Sr[i]|+|S|) $, where $ \Vr $ and $ \Sr[i] $ are the node set and snapshot of the inputed temp-multi-network, $ S $ is the summative snapshot of the inputed temp-multi-network. Our goal is to get the value of time complexity of Algorithm~\ref{alg:temp_cuboid_baseline} with corresponding $ \Gr_i'' $ of each $ A_i'' $ as the inputed temp-multi-network in $ O(1) $ time and then choose the lowest time complexity and the corresponding $ \Gr_i'' $. We can not stand a higher time complexity to compute the time complexity of Algorithm~\ref{alg:temp_cuboid_baseline}, which will make the process of choosing an ideal $ \Gr_i'' $ unworthy compared to choosing a arbitrary $ \Gr_i'' $. In time complexity of Algorithm~\ref{alg:temp_cuboid_baseline}, . For our basic Algorithm~\ref{alg:temp_cuboid_baseline}, if we maintain a prefix array $ pArray $ for each precomputed $ \Gr_i'' $ of $ A_i'' $, in which
%\begin{equation}
%	\label{eqt:pre_array_baseline}
% 	pArray[i]=\sum_{j=1}^{i}|\Sr''[j]|, i \leq t_{end},
%\end{equation}
%$ \Sr''[j] $ is the snapshot at timestamp $ j+t_{base} $ as defined in Definition~\ref{def:snapshot}, we can easily compute the value of accumulation part ($ \sum_{i=l-t_{base}}^{r-t_{base}}|\Sr[i]| $) in the time complexity of Algorithm~\ref{alg:temp_cuboid_baseline} by
%\begin{equation}
%	\label{eqt:pre_array_base_use}
%	pArray[r-t_{base}]-pArray[l-t_{base}-1].
%\end{equation}
%It's easy to update $ pArray $ and $ \Sr''[j] $ according to their definitions when adding new edges at the latest timestamp in each $ \Gr_i'' $ (we will talk about adding new edges into temporal graph cube in Section~\ref{subsec:growing}). As a result, we can compute the value of time complexity of Algorithm~\ref{alg:temp_cuboid_baseline} with different $ \Gr_i'' $ as the inputed temp-multi-network by $ |\Vr_i''|+pArray[r-t_{base}]-pArray[l-t_{base}-1]+|S_i''| $ in $ O(1) $ time. .[not accurate because of $ |S_i''| $]

After discussing temporal cuboid query in detail, we can clearly describe the support to OLAP operations in temporal graph cube, such as roll-up, drill-down and slice-and-dice. Suppose we have conducted a temporal cuboid query $ Q_0=(A_0,[l,r]) $. Roll-up means conducting another query $ Q_1=(A_1,[l,r]) $ where $ A_1 $ is an ancestor of $ A_0 $, so that we get a coarser resolution of summative information of temp-multi-network in the same time range. Drill-down is a contrary operation to obtain a finer summative information in the same time range by conducting $ Q_2=(A_2,[l,r]) $ where $ A_2 $ is a descendent of $ A_0 $. Slice-and-dice can be performed by selecting a subset of vertices in the vertex table of $ Q_0 $'s result and generating the induced network. For example, $ Q_0=((Profession),[2,5]) $ in temporal graph cube built on temp-multi-network in Figure~\ref{fig:example_dataset}, we can choose to build an induced network on the basis of the result of $ Q_0 $ with only ``teacher" and ``student" being selected to clearly show the transactions between ``teacher" and ``student" from day 2 to day 5.

We can also extend roll-up and drill-down according to the specified time ranges. A time related roll-up means conducting a query $ Q_1=(A_0,[l_1,r_1]) $ where $ l_1 \leq l $ and $ r_1 \geq r $, so that we can get a coarser resolution of summative information in time axis. On the contrary, time related drill-down means conducting a query $ Q_2=(A_0,[l_2,r_2]) $ where $ l_2 \geq l $ and $ r_2 \leq r $.

\stitle{Temporal crossboid.} The inputs of temporal crossboid query $ Q_{cross}=(A'_1,A'_2,[l,r]) $ in temporal graph cube contain two specified views $ A_1'=(A_{11}',A_{12}',...,A_{1n}') $, $ A_2'=(A_{21}',A_{22}',...,A_{2n}') $ and a time interval $ [l,r] $, where $ A_1' \neq A_2' $. The output is a static aggregate bipartite network with two types of aggregated vertices corresponding to aggregations $ A_1' $ and $ A_2' $ respectively. For example, the question in Example~\ref{exam:crossboid_query} can be represented by $ Q_{cross}=((Gender),(Profession),[2,3]) $.

We can conduct a temporal crossboid query directly on the original temp-multi-network. However, it's inefficient due to the size of the original temp-multi-network. We omit the algorithm to conduct a temporal crossboid query on the original temp-multi-network, but give a more efficient way to conduct the query.
\begin{definition}[Nearest Common Descendant]
	\label{def:ncd}
	Given two different views $ A_1' $ and $ A_2' $, the common descendant of $ A_1' $ and $ A_2' $, $ cd(A_1',A_2') $ is also a view in temporal graph cube, satisfying $ dim(A_1') \cup dim(A_2') \subseteq dim(cd(A_1',A_2')) $. The nearest common descendant of $ A_1' $ and $ A_2' $, $ ncd(A_1',A_2') $, is one of the common descendants satisfying $ dim(ncd(A_1',A_2'))=dim(A_1') \cup dim(A_2') $.
\end{definition}
\begin{theorem}
	\label{thm:ans_cross_by_ncd}
	Given a temporal crossboid query $ Q_{cross}=(A_1',A_2',[l,r]) $, $ Q_{cross} $ can be directly answered from the result of temporal cuboid query $ Q=(ncd(A_1',A_2'),[l,r]) $ where $ Q,Q_{cross} $ share the same aggregation function upon vertices and edges.
\end{theorem}
\begin{proof}
	Although the result of $ Q_{cross} $ is a bipartite network, we can treat the edges in it as directed edges and treat each edge in materialized views of the temporal graph cube and in the result of $ Q $ as two directed edges, so that we can use the similar idea from the previous proofs of Theorem~\ref{thm:cuboid_by_temp_aggre_network} and Theorem~\ref{thm:cuboid_by_small_temp_aggre_network} to prove this theorem. We omit the details of this proof due to the space limitation.
\end{proof}
With Theorem~\ref{thm:ans_cross_by_ncd}, a temporal crossboid query can be turned into a temporal cuboid query and we have discussed how to conduct a temporal cuboid query efficiently in \stitle{Temporal cuboid} section.
%\begin{theorem}
%	\label{thm:ans_cross_by_cuboid}
%	Given a temporal crossboid query $ Q_{cross}=(A_1',A_2',[l,r]) $, $ Q_{cross} $ can be directly answered from the result of a temporal cuboid query $ Q=(A_3',[l,r]) $ where $ dim(A_1') \cup dim(A_2') \subseteq dim(A_3') $ and $ Q,Q_{cross} $ share the same aggregation function upon vertices and edges.
%\end{theorem}
%\begin{proof}
%	todo
%\end{proof}
\begin{algorithm}[t]
	\scriptsize
	\caption{$\kw{Temporal Crossboid Query}$}
	\label{alg:temp_crossboid_on_cuboid}
	\KwIn{$ G=(V,E,B,W_{G}) $, result of $ Q=(ncd(A_1',A_2'),[l,r]) $; A temporal crossboid query $ Q_{cross}=(A_1',A_2',[l,r]) $; Aggregation functions $f_v$ and $f_e$}
	\KwOut{A bipartite graph $ G_B=(V_1,V_2,E_B,W_1,W_2,B_1,B_2) $}
	$ B_1 \gets dim(A_1') $, $ B_2 \gets dim(A_2') $\;
	Let $ h_1 $ be hash structure: $ \{B_1(u)|u \in V\} \rightarrow V_1 $, $ h_2 $ be hash structure: $ \{B_2(u)|u \in V\} \rightarrow V_2 $;

	\ForEach{$ u \in V $}{
		\If{$ h_1(B_1(u))=NULL $}{
			$ V_1 \gets V_1 \cup \{u\} $;
			$ W_1(u) \gets W_{G}(u) $;
			$ h_1(B_1(u)) \gets u $;
		}
		\Else{
			$ u' \gets h_1(B_1(u)) $;
			$ W_1(u') \gets f_v(W_1(u'),W(u)) $;
		}
		\If{$ h_2(B_2(u))=NULL $}{
			$ V_2 \gets V_2 \cup \{u\} $;
			$ W_2(u) \gets W_{G}(u) $;
			$ h_2(B_2(u)) \gets u $;
		}
		\Else{
			$ u'' \gets h_2(B_2(u)) $;
			$ W_2(u'') \gets f_v(W_2(u''),W(u)) $;
		}
	}
	$ EP \gets \{\} $;
	$ E_B \gets \{\} $;

	\ForEach{$ (u,v,a) \in E $}{
		$ u' \gets h_1(u) $;
		$ v'' \gets h_2(v) $;

		\If{$ (u',v'') \notin EP $}{
			$ EP \gets EP \cup \{(u',v'')\} $;
			$ E_B \gets E_B \cup \{(u',v'',a)\} $;
		}
		\Else{
			Let $ (u',v'',a') $ be an edge in $ E_B $ which has $ u',v'' $ as vertices\;
			$ E_B \gets E_B-\{(u',v'',a')\} $\;
			$ E_B \gets E_B \cup \{(u',v'',f_e(a',a))\} $;
		}
		$ v' \gets h_1(v) $;
		$ u'' \gets h_2(u) $;

		\If{$ (v',u'') \notin EP $}{
			$ EP \gets EP \cup \{(v',u'')\} $;
			$ E_B \gets E_B \cup \{(v',u'',a)\} $;
		}
		\Else{
			Let $ (v',u'',a'') $ be an edge in $ E_B $ which has $ v',u'' $ as vertices\;
			$ E_B \gets E_B-\{(v',u'',a'')\} $\;
			$ E_B \gets E_B \cup \{(v',u'',f_e(a'',a))\} $;
		}
	}
	\Return $ G_B=(V_1,V_2,E_B,W_1,W_2,B_1,B_2) $;
\end{algorithm}
Algorithm~\ref{alg:temp_crossboid_on_cuboid} is an algorithm to compute the result of $ Q_{cross}=(A_1',A_2',[l,r]) $ from the result of $ Q=(ncd(A_1',A_2'),[l,r]) $. Unlike Algorithm~\ref{alg:temp_cuboid_baseline}, we need aggregate each vertex on two different aggregations $ A_1' $ and $ A_2' $ (line~3-11). For each edge $ (u,v,a) $ in $ G $, we need to create or update two aggregated edges $ (u',v'',\ast) $ and $ (v',u'',\ast) $ (line~12-27), because the two directions of $ (u,v,a) $ represent two interactions of aggregated vertices respectively (note that we don't assign $ u' \leq v'' $ or $ v' \leq u'' $ in $ E_B $). The time complexity of Algorithm~\ref{alg:temp_crossboid_on_cuboid} is $ O(|V|+|E|) $. We can easily find that $ |V_1|+|V_2| \leq 2|V| $ and $ |E_B| \leq 2|E| $ in Algorithm~\ref{alg:temp_crossboid_on_cuboid}, so the space complexity of Algorithm~\ref{alg:temp_crossboid_on_cuboid} is also $ O(|V|+|E|) $.

\subsection{Growth of Network}\label{subsec:growing}
Compared to static network, temporal network records ongoing relationships between vertices, which means we should append new edges into temporal network. For example, the temp-multi-network in Figure~\ref{fig:example_dataset} should record new transactions between any two individuals in day 5 and in all following days. Specifically, suppose we need to add a new edge $ (u,v,t,a) $ with discrete and numeric attributes $ (A(u),W(u),A(v),W(v)) $ of two vertices, there are two situations: 1) $ t-t_{base}=t_{end} $ ($ t=t_{|\Tr|} $, i.e., $ t $ equals the latest timestamp in the temporal graph cube). 2) $ t-t_{base}>t_{end} $.

For the original temp-multi-network $ \Gr=(\Vr,\Sr,A,W) $, it's simple to add a new edge in both situations. In situation 1), we only need to add $ u,v $ into $ \Vr $, update $ A,W $ with $ (A(u),W(u),A(v),W(v)) $ if $ u $ and $ v $ are new vertices and then $ (u,v,a) $ will be added into $ \Sr[t_{end}] $. In situation 2), the only difference is that we should extend $ \Sr $ and create an empty snapshot $ \Sr[t-t_{base}] $, then we add $ (u,v,a) $ into $ \Sr[t-t_{base}] $. Finally we update $ t_{end} $ to $ t-t_{base} $.

For any precomputed view $ A' $ and its corresponding temporal aggregate network $ \Gr'=(\Vr',\Sr',B,W_{\Gr'}) $, if $ u $ is a new vertex in the original temp-multi-network, we need to create or update an aggreagted vertex $ u' $ in $ \Vr' $ according to $ h(B(u)) $, where $ h $ is the hash structure used in construction of $ \Gr' $ and update $ h $ if $ h(B(u))=NULL $ (line~3-6 in Algorithm~\ref{alg:view_construct_baseline}). We do the same for $ v $. Then we map $ (u,v,a) $ to $ (u',v',a) $ (line~11-12 in Algorithm~\ref{alg:view_construct_baseline}), which can be regarded as the new edge for $ \Gr' $. If situation 1) is met, we create or update aggregated edge $ (u',v',\ast) $ in $ \Sr'[t_{end}] $ with $ (u',v',a) $ as line~13-18 in Algorithm~\ref{alg:view_construct_baseline}. If situation 2) is met, we also extend $ \Sr' $ and create an empty snapshot $ \Sr'[t-t_{base}] $, then add $ (u',v',a) $ directly into $ \Sr'[t-t_{base}] $ and update $ t_{end} $ to $ t-t_{base} $ in $ \Gr' $.

\begin{figure}[t!]
	\begin{center}
		\begin{tabular}[t]{c}\hspace*{-0.3cm}
			\subfigure[{\scriptsize Original network}]{
				\includegraphics[width=0.33\columnwidth]{fig/example/add_edge_at_5_original.pdf}
			}
			\hspace*{-0.3cm}
			\subfigure[{\scriptsize (Gender)}]{
				\includegraphics[width=0.33\columnwidth]{fig/example/add_edge_at_5_gender_adjust.pdf}
			}
			\hspace*{-0.3cm}
			\subfigure[{\scriptsize (Gender, Profession)}]{
				\includegraphics[width=0.33\columnwidth]{fig/example/add_edge_at_5_gender_pro_adjust.pdf}
			}
		\end{tabular}
	\end{center}
	\vspace*{-0.4cm}
	\caption{Snapshots of temp-multi-network in memory at day 5 after new edge is added}
	\vspace*{-0.4cm}
	\label{fig:add_edge_at_5}
\end{figure}

\begin{figure}[t!]
	\begin{center}
		\begin{tabular}[t]{c}\hspace*{-0.3cm}
			\subfigure[{\scriptsize Original network}]{
				\includegraphics[width=0.33\columnwidth]{fig/example/add_edge_at_6_original.pdf}
			}
			\hspace*{-0.3cm}
			\subfigure[{\scriptsize (Gender)}]{
				\includegraphics[width=0.31\columnwidth]{fig/example/add_edge_at_6_gender_adjust.pdf}
			}
			\hspace*{-0.3cm}
			\subfigure[{\scriptsize (Gender, Profession)}]{
				\includegraphics[width=0.33\columnwidth]{fig/example/add_edge_at_6_gender_pro_adjust.pdf}
			}
		\end{tabular}
	\end{center}
	\vspace*{-0.4cm}
	\caption{Snapshots of temp-multi-network in memory at day 6 after new edge is added}
	\vspace*{-0.4cm}
	\label{fig:add_edge_at_6}
\end{figure}
\begin{example} \label{exam:network_growth}
	Suppose we need to add two new edges $ e_1=(v_2,v_3,5,10) $ and $ e_2=(v_6,v_9,6,5) $ successively where $ A(v_9)=(female,cn,student) $, $ W(v_9)=800 $ into temporal graph cube built on temp-multi-network in Figure~\ref{fig:example_dataset} and only two views are precomputed: $ (Gender,\ast,\ast) $ and $ (Gender,\ast,Profession) $. Figure~\ref{fig:add_edge_at_5} shows the snapshots at day 5 of all temp-multi-networks in the temporal graph cube after $ e_1 $ is added. Since $ v_2 $ and $ v_3 $ are both females, the self loop edge on ``female" in the 5-th snapshot in Figure~\ref{fig:aggregation_gender} should be updated to the corresponding self loop in Figure~\ref{fig:add_edge_at_5} (b). Similar update should be performed in the 5-th snapshot in Figure~\ref{fig:aggregation_gender_profession}, and the result is shown in Figure~\ref{fig:add_edge_at_5} (c).

	When adding $ e_2 $, we need to create new snapshot in each temp-multi-network of the temporal graph cube because $ e_2 $ leads to a new timestamp day 6. For the original network, we create a new snapshot at day 6 as Figure~\ref{fig:add_edge_at_6} (a) with only $ e_2 $ being stored after all attributes of $ v_9 $ are added into the vertex table of the original temp-multi-network. For other temp-multi-networks of precomputed views, we should also create new snapshots at day 6, holding the aggregated edges of $ e_2 $ as in Figure~\ref{fig:add_edge_at_6} (b) and (c) after the aggregated vertices of $ v_9 $ are added or updated.
\end{example}

\section{Indexes} \label{sec:indexes}
In this section, we mainly discuss how to build indexes in temporal graph cube to accelerate temporal cuboid query (temporal crossboid query can be turned into temporal cuboid query). In Section~\ref{subsec:queries}, given a temporal cuboid query $ Q=(A',[l,r]) $, we merge snapshots of a certain temp-multi-network in time range $ [l,r] $  (i.e., line~8-15 of Algorithm~\ref{alg:temp_cuboid_baseline}), whether $ A' $ is precomputed or not. Merging snapshots in time ranges is similar to common range query problem: answering plenty of online queries $ q=(func,[l,r]) $ on a numeric array $ Arr $, where $ func $ could be SUM($\ast$), AVERAGE($\ast$), MAX($\ast$) or MIN($\ast$), specifying the needed statistical result of values in range $ [l,r] $ of $ Arr $. We can regard merging snapshots in snapshot array as range query problem on snapshot array.

\begin{figure}[t!]
	\begin{center}
		\begin{tabular}[t]{c}\hspace*{-0.3cm}
			\subfigure[{\scriptsize A numeric array}]{
				\includegraphics[width=0.85\columnwidth]{fig/example/example_numeric_array_adjust.pdf}
			}
			\hspace*{-0.3cm}
			\\
			\subfigure[{\scriptsize A snapshot array extracted from Figure~\ref{fig:aggregation_gender_profession}}]{
				\includegraphics[width=0.95\columnwidth]{fig/example/example_snapshot_array.pdf}
			}
			\hspace*{-0.3cm}
			\\
			\subfigure[{\scriptsize Result of range query $ Q $ in Example~\ref{exam:range_query_num_snap_array}}]{
				\includegraphics[width=0.45\columnwidth]{fig/example/example_range_query_snapshot.pdf}
			}
		\end{tabular}
	\end{center}
	\vspace*{-0.4cm}
	\caption{Range query on numeric and snapshot array}
	\vspace*{-0.4cm}
	\label{fig:range_query_num_snap_array}
\end{figure}
\begin{example} \label{exam:range_query_num_snap_array}
	Figure~\ref{fig:range_query_num_snap_array} (a) shows a numeric array of length 5. Figure~\ref{fig:range_query_num_snap_array} (b) is the snapshot array of temporal aggregate network of precomputed view $ (Gender,\ast,Profession) $ in Figure~\ref{fig:aggregation_gender_profession} but we renumber all vertices and omit the isolated vertices in each snapshot for simplicity. For example, range query on numeric array in Figure~\ref{fig:range_query_num_snap_array} (a) can be $ q=(SUM(\ast),[2,4]) $, which queries the amount of values in $ [2,4] $ of the numeric array. Range query on snapshot array in Figure~\ref{fig:range_query_num_snap_array} (b) can be $ Q=((Gender,\ast,Profession),[2,4]) $ with SUM($\ast$) specified upon edges. The result of $ Q $ is shown in Figure~\ref{fig:range_query_num_snap_array} (c).
\end{example}

There are mainly two differences between range query on snapshot array and numeric array:
\begin{itemize}[leftmargin=*]
	\item We should merge both the structural information of snapshots and the numeric attributes of edges in snapshots in certain ranges. Range query on snapshot array can be regarded as range query on set array, in which we calculate the union of sets in specified range and maintain the statistical results of numeric attributes on all elements (edges).

	\item Adding (or appending) edges should be considered. In range query problem on numeric array, modifying the value at an arbitrary place in the array is sometimes being considered, but the size of the array is hardly changed. However in temp-multi-network, with the time goes on, more edges will be added into the temp-multi-network, while the past edges are seldom to be modified.
\end{itemize}

There are many existing methods to build indexes to accelerate range query on numeric array. In this paper we implement some representative indexes to accelerate range query on snapshot array and extend them to support self maintenance when new edges are added. Specifically, we divide our problems into two kinds according to aggregation function specified upon edges $ f_e $: 1) $ f_e = $ COUNT($\ast$) or SUM($\ast$). 2) $ f_e= $ MAX($\ast$) or MIN($\ast$). In range query problem on numeric array, if $ func= $ SUM($\ast$) (COUNT($\ast$) is meaningless most of the time) in a query and there is no modification at arbitrary place, we can maintain a simple prefix array $ preArr $, in which
\begin{equation}
	\label{eqt:prearr_range_query}
	preArr[i]=\sum_{j=0}^{j \leq i}Arr[j].
\end{equation}
A query $ q=(func,[l,r]) $ can be easily answered by
\begin{equation}
	\label{eqt:prearr_ans_range_query}
	preArr[r]-preArr[l-1].
\end{equation}
However, if $ func= $ MAX($\ast$) or MIN($\ast$) and we still maintain each $ preArr[i] $ as the maximum (or minimum) of $ Arr $ in $ [0,i] $, we can not tell the maximum (or minimum) in $ [l,r] $ of $ Arr $ if $ preArr[r]=preArr[l-1] $.

It's the same with our problem because range query on numeric array can be regarded as a special case of our problem if each snapshot in snapshot array has only one edge and all edges of those snapshots share the same vertices. As a result, in Section~\ref{subsec:prefix_array}, we briefly discuss how to use prefix array to accelerate temporal cuboid queries when $ f_e $ equals COUNT($\ast$) or SUM($\ast$) and maintain prefix array while new edges are added into the temporal graph cube. In Section~\ref{subsec:sparse_table} and Section~\ref{subsec:segment_tree}, we discuss two different methods to accelerate queries when $ f_e $ equals MAX($\ast$) or MIN($\ast$) and their maintenance while new edges added.

\subsection{Prefix Array} \label{subsec:prefix_array}
Before discussing prefix array, we need to explain that in practical implementation, to support all kinds of aggregation functions we maintain all kinds of statistical results at the same time, i.e., in all snapshots of all temporal aggregate networks of precomputed views, each edge is in form of $ (u,v,c,a,max,min) $ where $ c,a $ are count of edges and amount of attributes respectively.

\stitle{Build.} We can build a prefix array for each precomputed view and $ A_{base} $. The elements in prefix array is still snapshots, but edges in these snapshots are in form of $ (u,v,c,a) $, since we only maintain counts of edges and amounts of numeric attributes of edges using prefix array.
\begin{algorithm}[t]
	\scriptsize
	\caption{$\kw{Build Prefix Array}$}
	\label{alg:build_prefix_array}
	\KwIn{A temp-multi-network $ \Gr=(\Vr,\Sr,A,W) $}
	\KwOut{Prefix Array $ preArray $ for $ \Gr $}
	Let $ preArray $ be an array initialized with $ t_{end}+1 $ empty snapshots;

	\ForEach{$ i \in [1,t_{end}] $}{
		$ preArray[i] \gets \Sr[i] $;
	}

	\ForEach{$ i \in [2,t_{end}] $}{
		$ \kw{MergeSnapshotSum}(preArray[i],preArray[i-1]) $;
	}
	\Return $ preArray $;

	\vspace*{0.2cm}
	{\bf Procedure} {$\kw{MergeSnapshotSum}(S_a,S_b)$}

	\ForEach{$ e_b \in S_b $}{
		\If{$\exists e_a \in S_a, e_a.u=e_b.u \wedge e_a.v=e_b.v $}{
			$ e_a.c \gets e_a.c+e_b.c $\;
			$ e_a.a \gets e_a.a+e_b.a $;
		}
		\Else{
			$ S_a \gets S_a \cup \{(e_b.u,e_b.v,e_b.c,e_b.a)\} $;
		}
	}
\end{algorithm}

We construct prefix array in Algorithm~\ref{alg:build_prefix_array}. Line~2-5 is the basic process of building a prefix array. The key part in Algorithm~\ref{alg:build_prefix_array} is the \kw{MergeSnapshotSum} procedure, which merges a snapshot $ S_b $ into another snapshot $ S_a $. In detail, for each edge $ e_b $ in $ S_b $, we add $ e_b $ into $ S_a $ (line~13) or update $ e_a $ in $ S_a $ which shares the same vertices with $ e_b $ (line~10-11).

Suppose the summative snapshot of $ \Gr $ is $ S $, the time complexity and space complexity of Algorithm~\ref{alg:build_prefix_array} are both $ O(t_{end}\times |S|) $.

\stitle{Query.} It's easy to use prefix array to answer $ Q=(A',[l,r]) $ when $ f_e $ is COUNT($\ast$) or SUM($\ast$). Suppose conducting $ Q $ we need merge snapshots in $ [l,r] $ of a temp-multi-network $ \Gr $ ($ \Gr $ may not be the corresponding temporal aggregate network of $ A' $, the reason can be found in Section~\ref{subsec:queries}), and prefix array built for $ \Gr $ is $ preArray $, then the result of merging snapshots of $ \Gr $ can be easily obtained by Algorithm~\ref{alg:prefix_array_query}. If $ A' $ is precomputed, then $ \Gr=\Gr' $ and $ S_E $ (the output of Algorithm~\ref{alg:prefix_array_query}) is the network structure of $ Q $'s result, otherwise the same process in line~17-25 of Algorithm~\ref{alg:temp_cuboid_baseline} should be performed.
\begin{algorithm}[t]
	\scriptsize
	\caption{$\kw{Prefix Array Query}$}
	\label{alg:prefix_array_query}
	\KwIn{Prefix array $ preArray $ for $ \Gr $; Biased time range $ [l,r] $}
	\KwOut{A snapshot $ S_E $, the result of merging snapshots of $ \Gr $ in $ [l,r] $ }
	$ S_E \gets \{\} $;
	$ l \gets l-1 $;

	\ForEach{$ e_r \in preArray[r] $}{
		\If{$\exists e_l \in preArray[l], e_r.u=e_l.u \wedge e_r.v=e_l.v $}{
			\lIf{$ e_r.c>e_l.c $}{$ S_E \gets S_E \cup \{(e_r.u,e_r.v,e_r.c-e_l.c,e_r.a-e_l.a)\} $}
		}
		\Else{
			$ S_E \gets S_E \cup \{(e_r.u,e_r.v,e_r.c,e_r.a)\} $;
		}
	}
	\Return $ S_E $;
\end{algorithm}
It's clear that the time complextiy of Algorithm~\ref{alg:prefix_array_query} is $ O(|preArray[r]|) $.

\stitle{Update.} We have discussed that there are two situations when adding a new edge $ (u,v,t,a) $: 1) $ t-t_{base}=t_{end} $. 2) $ t-t_{base}>t_{end} $. We directly give the updating algorithm Algorithm~\ref{alg:prefix_array_update}.
\begin{algorithm}[t]
	\scriptsize
	\caption{$\kw{Prefix Array Update}$}
	\label{alg:prefix_array_update}
	\KwIn{Prefix array $ preArray $ for $ \Gr $; Mapped edge $ (u',v',a) $ of new edge $ e=(u,v,t,a) $ used to update $ \Gr $; $ t_{end} $ of $ \Gr $ before $ \Gr $ was updated}
	\KwOut{Updated $ preArray $}
	\If{$ t-t_{base}>t_{end} $}{
		Append $ (t-t_{base}-t_{end}) $ $ preArray[t_{end}] $s to $ preArray $;
	}
	\If{$ \exists e_t \in preArray[t-t_{base}],e_t.u=u' \wedge e_t.v=v' $}{
		$ e_t.c \gets e_t.c+1 $;
		$ e_t.a \gets e_t.a+a $;
	}
	\Else{
		$ preArray[t-t_{base}] \gets preArray[t-t_{base}] \cup \{(u',v',1,a)\} $;
	}
	\Return $ preArray $;
\end{algorithm}

In Algorithm~\ref{alg:prefix_array_update}, the mapped edge $ (u',v',a) $ in the input is used to update $ \Gr $ as we discussed in Section~\ref{subsec:growing}. If $ \Gr $ is the original temp-multi-network, $ (u',v',a)=(u,v,a) $. $ t_{end} $ is the biased timestamp of the latest timestamp in $ \Gr $ before $ \Gr $ was updated. If condition in line~1 is true, then the second situation of adding a new edge is met, and we should extend $ preArray $ to a new length by appending $ (t-t_{base}-t_{end}) $ $ preArray[t_{end}] $s to $ preArray $ (line~2). Whether condition in line~1 is true or false, the last step is updating $ preArray[t-t_{base}] $ with $ (u',v',a) $ (line~3-6).

The time complexity of Algorithm~\ref{alg:prefix_array_update} still depends on situations of adding a new edge. If the first situation is met, the time complexity is $ O(1) $. If the second situation is met, the time complexity is $ O((t-t_{base}-t_{end})\times |preArray[t_{end}]|) $.
\subsection{Sparse Table} \label{subsec:sparse_table}
Sparse table (STable) is an efficient data structure to accelerate range query on numeric array $ Arr $ with $ O(1) $ time complexity and $ O(nlog(n)) $ space complexity ($ n $ is the length of $ Arr $) only when $ func $ is MAX($\ast$) or MIN($\ast$) in $ q $. If $ func $ is SUM($\ast$), time complexity of conducting $ q $ becomes $ O(log(n)) $. We briefly explain the principle of STable bellow.
\begin{itemize}[leftmargin=*]
	\item For a numeric array $ Arr $ of length $ n $, STable maintains a two-dimensional array $ f $. $ f[i][j] $ represents the maximum (or minimum) in range $ [i-2^j+1,i] $ of $ Arr $.

	\item For each query $ q=(func,[l,r]) $, let $ len=r-l+1 $, if $ log(len) $ is an integer, then the answer of $ q $ is $ f[r][log(len)] $. Otherwise, let $ m=\lfloor log(len) \rfloor $, the answer of $ q $ is $ func(f[l+2^m-1][m],f[r][m]) $.
\end{itemize}
When $ log(len) $ is not an integer, we can not use a single element in $ f[r] $ to answer $ q $, but we can use two elements in $ f[l+2^m-1] $ and $ f[r] $, representing two overlapping ranges to answer $ q $. If $ func $ is SUM($\ast$), we can not use the same method to answer $ q $, because in this case, values in overlapping area will be added twice. It doesn't matter if $ func $ is MAX($\ast$) (or MIN($\ast$)), because MAX($ x,x $) still equals $ x $.

For a snapshot $ S $, we can also get MAX($ S,S $) $ =S $, because numeric value $ x $ on each edge of $ S $ still satisfies MAX($ x,x $) $ =x $. Note that the execution of MAX($\ast$) when the parameters are snapshots is similar to \kw{MergeSnapshotExtre} in Algorithm~\ref{alg:stable_build}. Next, we discuss how to extend STable to solve our problem.

First, we switch all elements in $ f $ to snapshots. We first consider the case where $ f_e= $ MAX($\ast$). For a temp-multi-network $ \Gr $ and its snapshot array $ \Sr $, we have
\begin{equation}
	\label{eqt:stable_elements}
	\begin{split}
		f[i][j]=MAX(\Sr[i-2^j+1],\\\Sr[i-2^j+2],...,\Sr[i]).
	\end{split}
\end{equation}
That is, $ f[i][j] $ is the MAX($\ast$) result of snapshots of $ \Sr $ in time range $ [i-2^j+1,i] $.

\stitle{Build.} Building a STable relies on the following equation:
\begin{equation}
	\label{eqt:build_stable}
	f[i][j]=MAX(f[i][j-1],f[i-2^{j-1}][j-1]).
\end{equation}
To support MIN($\ast$), we use edges in form of $ (u,v,max,min) $ in snapshots of $ f $.
\begin{algorithm}[t]
	\scriptsize
	\caption{$\kw{STable Build}$}
	\label{alg:stable_build}
	\KwIn{A temp-multi-network $ \Gr=(\Vr,\Sr,A,W) $}
	\KwOut{STable $ f $ for $ \Gr $}
	Let $ f $ be an array initialized with $ t_{end}+1 $ empty arrays\;
	\ForEach{$ i \in [1,t_{end}] $}{
		$ \kw{BuildSinglePlace}(i,f,\Sr) $;
	}
	\Return $ f $;

	\vspace*{0.2cm}
	{\bf Procedure} {$\kw{MergeSnapshotExtre}(S_a,S_b)$}

	\ForEach{$ e_b \in S_b $}{
		\If{$\exists e_a \in S_a, e_a.u=e_b.u \wedge e_a.v=e_b.v $}{
			$ e_a.max \gets MAX(e_a.max,e_b.max) $\;
			$ e_a.min \gets MIN(e_a.min,e_b.min) $;
		}
		\Else{
			$ S_a \gets S_a \cup \{(e_b.u,e_b.v,e_b.max,e_b.min)\} $;
		}
	}

	\vspace*{0.2cm}
	{\bf Procedure} {$\kw{BuildSinglePlace}(pos,f,snapArr)$}

	$ f[pos].append(snapArr[pos]) $\;
	$ range \gets 2 $;
	$ i \gets 1 $\;
	\While{$ pos-range+1 \geq 1 $}{
		$ halfRange \gets range>>1 $;
		$ prePos \gets pos-halfRange $\;
		$ f[pos].append(f[pos][i-1]) $\;
		$ \kw{MergeSnapshotExtre}(f[pos][i],f[prePos][i-1]) $\;
		$ range \gets range<<1 $;
		$ i=i+1 $;
	}
\end{algorithm}
Algorithm~\ref{alg:stable_build} is used to build STable for every temp-multi-network.

In line~3 of Algorithm~\ref{alg:stable_build}, \kw{BuildSinglePlace} is executed to build each $ f[i],i\in [1,t_{end}] $. \kw{BuildSinglePlace} procedure is designed based on Equation~\ref{eqt:build_stable} as line~14-19. The key part of Algorithm~\ref{alg:stable_build} is the \kw{MergeSnapshotExtre} procedure, which conducts MAX($\ast$) for snapshots. The time complexity and space complexity of Algorithm~\ref{alg:stable_build} are both $ O(t_{end}\times log(t_{end})|S|) $ where $ S $ is the summative snapshot of the inputed temp-multi-network.

\stitle{Query.} Algorithm~\ref{alg:stable_query} conducts the range query on snapshot array using STable.
\begin{algorithm}[t]
	\scriptsize
	\caption{$\kw{STable Query}$}
	\label{alg:stable_query}
	\KwIn{STable $ STable $ for $ \Gr $; Biased time range $ [l,r] $}
	\KwOut{A snapshot $ S_E $, the result of merging snapshots of $ \Gr $ in $ [l,r] $}
	$ S_E \gets \{\} $;
	$ range \gets r-l+1 $\;
	$ f \gets STable.f $\;
	\If{$ log(range) $ is an integer}{
		$ S_E \gets f[r][log(range)] $;
	}
	\Else{
		$ logVal \gets \lfloor log(range) \rfloor $\;
		$ S_E \gets f[r][logVal] $\;
		$ powerVal \gets 1<<logVal $\;
		$ \kw{MergeSnapshotExtre}(S_E,f[l+powerVal-1][logVal]) $;
	}

	\Return $ S_E $;
\end{algorithm}
The idea of Algorithm~\ref{alg:stable_query} is similar to algorithms conducting queries on numeric arrays using STable. The only difference is the use of \kw{MergeSnapshotExtre}. The time complexity of Algorithm~\ref{alg:stable_query} is $ O(|S|) $

\stitle{Update.} It is generally considered that STable doesn't support to be updated when modifying at arbitrary places of $ Arr $ in range query problem on numeric array because of the low efficiency. However, in our problem, we only need to modify snapshot array at the last place (i.e., adding a new edge to the last snapshot) or create a new snapshot holding the new edge at the end of the snapshot array. It's possible to update STable in our problem.
\begin{algorithm}[t]
	\scriptsize
	\caption{$\kw{STable Update}$}
	\label{alg:stable_update}
	\KwIn{STable $ STable $ for $ \Gr=(\Vr,\Sr,A,W) $; Mapped edge $ (u',v',a) $ of new edge $ e=(u,v,t,a) $ used to update $ \Gr $; $ t_{end} $ of $ \Gr $ before $ \Gr $ was updated}
	\KwOut{Updated $ STable $}
	$ f \gets STable.f $\;
	\If{$ t-t_{base}>t_{end} $}{
		Append $ (t-t_{base}-t_{end}) $ empty arrays to $ f $;

		\ForEach{$ i \in [t_{end}+1,t-t_{base}] $}{
			$ \kw{BuildSinglePlace}(i,f,\Sr) $;
		}
	}
	\Else{
		\ForEach{$ S \in f[t_{end}] $}{
			$ \kw{MergeSnapshotExtre}(S,\{(u',v',a,a)\}) $;
		}
	}
	\Return $ STable $;
\end{algorithm}

We use Algorithm~\ref{alg:stable_update} to update STable. For the second situation of adding new edges, we only need to append $ (t-t_{base}-t_{end}) $ empty 1-D arrays to $ f $ and execute \kw{BuildSinglePlace} at extra places of $ f $ (line~3-5), because the update of STable is later than the update of $ \Gr $ the STable built on, and the process in line~3-5 of Algorithm~\ref{alg:stable_update} can be regarded as a continuing process of line~2-3 in Algorithm~\ref{alg:stable_build}.

The time complexity of Algorithm~\ref{alg:stable_update} still depends on situations of adding a new edge. If situation 1) is met, we should update each snapshot in $ STable.f[t_{end}] $ with mapped edge $ (u',v',a,a) $ (line~7-8), so the time complexity is $ O(log(t_{end})) $. If situation 2) is met, we should conduct \kw{BuildSinglePlace} for $ STable.f[i],i \in [t_{end}+1,t-t_{base}] $, so the time complexity is $ O((t-t_{base}-t_{end})\times log(t-t_{base})|S|) $, $ S $ is the summative snapshot of $ \Gr $ which STable is built on.

\subsection{Segment Tree} \label{subsec:segment_tree}
Segment tree (STree) is an efficient data structure to accelerate range query on numeric array, and it supports all kinds of statistical functions mentioned above.
\begin{figure}[t]
	\begin{center}
		\includegraphics{fig/example/example_segtree.pdf}
	\end{center}
	\vspace*{-0.4cm}
	\caption{An example of STree}
	\vspace*{-0.4cm}
	\label{fig:exam_segtree}
\end{figure}
For example, Figure~\ref{fig:exam_segtree} shows the structure of STree built for numeric array in Figure~\ref{fig:range_query_num_snap_array} (a). In this section we mainly discuss how to extend STree to support range query on snapshot array when $ f_e $ is MAX($\ast$) or MIN($\ast$) and the maintanance of STree when adding new edges.

We first briefly discribe the basic properties of STree.
\begin{enumerate}
	\item STree is a full binary tree, maintaining values for an array. If the length of the array is $ N $, then the height of STree is $ \lceil log(N) \rceil+1 $.
	\item Each node in STree maintains statistical result of values in a sub-range of the array. The root of STree maintains the whole range of the array, i.e., $ [1,N] $. The left child and the right child maintain range $ [1,\lfloor(1+N)/2\rfloor] $ and range $ [\lfloor(1+N)/2\rfloor+1,N] $ respectively, which split $ [1,N] $ equally. Both left child and right child can be regarded as roots of sub-STrees and they both have their own childs sharing their ranges equally. The above process for a node ends when its range can not be divided (length equals $ 1 $).
	\item The time complexity of building a STree is $ O(N) $. The space complexity of STree is still $ O(N) $. The time complexity of conducting a query $ q=(func,[l,r]) $ is $ O(log(N)) $.
\end{enumerate}
Differences of our problem are: 1, elements in the array and statistical results in nodes of STree are not numeric values but snapshots, with edges in form of $ (u,v,max,min) $. 2, the range STree maintains should be expanded because of the new added edges. Merging statistical information of nodes is a basic and frequent operation in both building and querying of STree, in which two statistical results of two nodes produce a new result. For the first difference, all we need to do is replacing the original operation on numeric values with \kw{MergeSnapshotExtre} procedure.

We will talk about the second difference in \stitle{Update.} section for STree. Biased timestamp is not required in STree, since timestamps are not used to access nodes in STree.

\stitle{Build.} In this section we omit the algorithm to build STree for snapshot array, because we only need to replace merging operation of two numeric values in algorithm building STree for numeric array with \kw{MergeSnapshotExtre} procedure. Suppose we build a STree $ STree $ for snapshot array in Figure~\ref{fig:range_query_num_snap_array} (b), the structure of $ STree $ can also be shown by Figure~\ref{fig:exam_segtree}, but each node of $ STree $ carries a snapshot.
\begin{figure}[t]
	\begin{center}
		\includegraphics{fig/example/example_stree_node_1_3.pdf}
	\end{center}
	\vspace*{-0.6cm}
	\caption{Snapshot carried by node maintaining range $ [1,3] $ in Figure~\ref{fig:exam_segtree}}
	\vspace*{-0.2cm}
	\label{fig:exam_stree_node}
\end{figure}
For example, node maintaining range $ [1,3] $ in Figure~\ref{fig:exam_segtree} carries the snapshot shown in Figure~\ref{fig:exam_stree_node}.

As we have mentioned before, time complexity of building STree for numeric array is $ O(N) $. Since \kw{MergeSnapshotExtre} could be called in building each node, the time complexity of building STree for snapshot array of length $ N $ is $ O(N|S|) $, where $ S $ is the summative snapshot of the snapshot array. The sapce consumed by STree built for the snapshot array is also $ O(N|S|) $. Note that each node $ node $ in STree is a struct, where $ node.lchild $ and $ node.rchild $ are left child and right child of $ node $ respectively and $ [node.timel,node.timer] $ is the range which $ node $ maintains. For leaves of STree, $ lchild $ and $ rchild $ are $ NULL $. Suppose the snapshot array STree built on is $ \Sr $, $ node.snapshot $ can be regarded as the summative snapshot of snapshot sub-array $ [\Sr[node.timel-t_{base}],...,\Sr[node.timer-t_{base}]] $. STree $ STree $ is also a struct, where $ STree.root $ is the root of $ STree $.

\stitle{Query.} Algorithm~\ref{alg:stree_query} conducts the range query on snapshot array using STree, and it shares the same framework with original query algorithm on STree but uses \kw{MergeSnapshotExtre} to merge snapshots. Merging snapshots only happens when the current split range $ [timel,timer] $ equals the range held by the current node $ treeNode $ (line~2 in Algorithm~\ref{alg:stree_query}). In the worst case, it happens at each depth of STree, so the time complexity of Algorithm~\ref{alg:stree_query} is $ O(log(N)|S|) $, where $ S $ is the summative snapshot.

For example, suppose we conduct a query $ Q=(\ast,[2,5]) $ on snapshot array in Figure~\ref{fig:range_query_num_snap_array} (b) (the view in $ Q $ is omitted), we have to merge snapshots in nodes with range $ [2,2] $, $ [3,3] $ and $ [4,5] $ as in grey nodes in Figure~\ref{fig:exam_segtree}.
\begin{algorithm}[t]
	\scriptsize
	\caption{$\kw{STreeQuery(treeNode,timel,timer,ans)}$}
	\label{alg:stree_query}
	\KwIn{$ treeNode $, node to be visited; Time range $ [timel,timer] $; $ ans $, a snapshot holding the current result}
	\lIf{$ treeNode=NULL $}{\Return}
	\If{$ timel=treeNode.timel \wedge timer=treeNode.timer $}{
		$\kw{MergeSnapshotExtre}(ans,treeNode.snapshot)$\;
		\Return;
	}
	$ mid \gets (treeNode.timel+treeNode.timer)/2 $\;
	\lIf{$ timel > mid $}{$ STreeQuery(treeNode.rchild,timel,timer,ans) $}
	\lElseIf{$ timer \leq mid $}{$ STreeQuery(treeNode.lchild,timel,timer,ans) $}
	\Else{
		$ STreeQuery(treeNode.lchild,timel,mid,ans) $\;
		$ STreeQuery(treeNode.rchild,mid+1,timer,ans) $;
	}
	\Return;
\end{algorithm}

\stitle{Update.} The biggest timestamp that STree maintains is $ STree.root.timer $, so the two situations in updating STree when adding new edge $ (u,v,t,a) $ are: 1) $ t \leq STree.root.timer $. 2) $ t>STree.root.timer $. If situation 2) is met, we need to expand the range STree maintains by creating new nodes maintaining larger range and switch the root of STree to one of the new nodes.

In detail, suppose $ l=STree.root.timel,r=STree.root.timer $, if $ t>r $, we first create a new node $ newNode $ with range $ [l,2r-l+1] $, which is 2 times the length of $ [l,r] $, so that $ STree.root $ can be a valid left child of $ newNode $. Then we let $ STree.root $ be the left child of $ newNode $. Finally we let $ newNode $ be the new root of $ STree $. If $ [l,2r-l+1] $ still can not cover $ t $, then we continue the process above till $ t $ is coverd. The updated $ STree $ is not a fully binary tree, since there is only left subtree for the new root which violates the basic properties of STree. However, it can be seen as part of a standard STree, which may be completed by continuously added new edges.
\begin{figure}[t]
	\begin{center}
		\begin{tabular}[t]{c}\hspace*{-0.3cm}
			\subfigure[{\scriptsize Expanded STree waiting for the new edge}]{
				\includegraphics[width=0.95\columnwidth]{fig/example/example_segtree_expand.pdf}
			}
			\hspace*{-0.3cm}
			\\
			\subfigure[{\scriptsize Expanded STree being inserted with a new edge}]{
				\includegraphics[width=0.85\columnwidth]{fig/example/example_segtree_expand_inserted.pdf}
			}
		\end{tabular}
	\end{center}
	\vspace*{-0.4cm}
	\caption{Update STree when new edges in the second situation are added}
	\vspace*{-0.4cm}
	\label{fig:exam_segtree_expand}
\end{figure}
\begin{example} \label{exam:stree_expand}
	Suppose we have built a STree $ STree $ for snapshot array in Figure~\ref{fig:range_query_num_snap_array} (b) and $ STree.root.timer=5 $. If we want to add the new edge $ e_2 $ with timestamp 6 in Example~\ref{exam:network_growth}, we should update $ STree $ so that it can cover $ [1,6] $. The updated $ STree $ is shown in Figure~\ref{fig:exam_segtree_expand} (a), where we create a new node $ newNode $ with range $ [1,10] $ and let $ STree.root $ be the $ lchild $ of $ newNode $. Finally we assign $ newNode $ to be the new root of $ STree $. The updated $ STree $ without the nodes presented by dotted rectangles in Figure~\ref{fig:exam_segtree_expand} (a) is not a complete STree.

	As in Figure~\ref{fig:exam_segtree_expand} (b) (we omit the left subtree of the root for simplicity), only nodes on path from the root to leaf with range $ [6,6] $ are needed to maintain the aggregated edge of $ e_2 $, so we only need to create or update nodes on that path.
\end{example}

\begin{algorithm}[t]
	\scriptsize
	\caption{$\kw{UpdateSTree}$}
	\label{alg:update_stree}
	\KwIn{STree $ STree $ built for $ \Gr $; Mapped edge $ (u',v',a) $ of new edge $ e=(u,v,t,a) $ used to update $ \Gr $;}
	\KwOut{Updated STree $ STree $}
	\While{$ STree.root.timer<t $}{
		$ newRoot \gets $ a new empty STree node\;
		$ newRoot.timel \gets STree.root.timel $\;
		$ newRoot.timer \gets 2 \times STree.root.timer+1-STree.root.timel $\;
		$ newRoot.lchild \gets STree.root $\;
		$ newRoot.snapshot \gets STree.root.snapshot $\;
		$ STree.root \gets newRoot $;
	}
	$\kw{STreeInsertEdge}(STree.root,(u',v',t,a))$;

	\Return $ STree $;

	\vspace*{0.2cm}
	{\bf Procedure} {$\kw{STreeInsertEdge}(treeNode,e'=(u',v',t,a))$}

	$ \kw{MergeSnapshotExtre}(treeNode.snapshot,\{(u',v',a,a)\}) $;

	\lIf{$ treeNode.timel=treeNode.timer $}{\Return}
	$ mid \gets \lfloor(treeNode.timel+treeNode.timer)/2\rfloor $\;
	\If{$ t \leq mid $}{
		\If{$ treeNode.lchild=NULL $}{
			$ newNode \gets $ a new empty STree node\;
			$ treeNode.lchild \gets newNode $\;
			$ newNode.timel \gets treeNode.timel $\;
			$ newNode.timer \gets mid $;
		}
		$\kw{STreeInsertEdge}(treeNode.lchild,e')$;
	}
	\Else{
		\If{$ treeNode.rchild=NULL $}{
			$ newNode \gets $ a new empty STree node\;
			$ treeNode.rchild \gets newNode $\;
			$ newNode.timel \gets mid+1 $\;
			$ newNode.timer \gets treeNode.timer $;
		}
		$ \kw{STreeInsertEdge}(treeNode.rchild,e') $;
	}
\end{algorithm}

Algorithm~\ref{alg:update_stree} is used to update STree. In line~1, the condition in \textbf{while} is true means that the current root of $ STree $ can not cover $ t $. So we need to create a new root, which has a double size of range compared to the current root (line~3-4). In line~6, we directly copy the snapshot of the current root to the new root, since there is no possible right child for the new root before new edge is added. If the range of the candidate new root can not cover $ t $ (it hardly happens because of the density of timestamps in real-word large scale temporal network), then the loop in line~1 continues. When $ STree.root.timer \geq t $ is true, we can use \kw{STreeInsertEdge} procedure to update STree. The framework of \kw{STreeInsertEdge} is similar to procedure used to update STree in range query problem on numeric array. In line~11, we update the snapshot of $ treeNode $ with the mapped edge, since $ t $ is in $ [treeNode.timel,treeNode.timer] $. In line~15 and line~22, we need to test whether the left child or the right child exists or not and create left child (line~16-19) or right child (line~23-26). We have discussed the reason before, that the STree is not complete after being expanded to maintain a larger range.

The time complexity of Algorithm~\ref{alg:update_stree} contains two parts: creating new root in line~1-7 and inserting new edge using \kw{STreeInsertEdge} procedure. Suppose the length of range maintained by the root of STree is $ len_1=STree.root.timer-STree.root.timel+1 $ before new edge is added, and the length of new range of timestamps in $ \Gr $ becomes $ len_2=t-STree.root.timel+1 $ after $ \Gr $ was updated. To cover $ len_2 $, we need repeat the loop in line~1 at least $ x $ times, and we have
\begin{equation}
	\label{eqt:new_root_create_times}
	\begin{split}
		len_1*2^x &\geq len_2\\
		x &\geq log(\frac{len_2}{len_1}).
	\end{split}
\end{equation}
For each loop, the cost mainly comes from line~6. Suppose the snapshot maintained in root is $ S_{root} $ before new edge is added, the time complexity of creating new root is $ O(log(\frac{len_2}{len_1})\times |S_{root}|) $. In practical, $ len_2>len_1 $ is hardly true because $ len_1 $, the length of range maintained by root, grows exponentially (line~3-4) but $ len_2 $ grows linearly since timestamps in real-world temporal graph are dense. In \kw{STreeInsertEdge} procedure, let the length of range in the final new root is $ len $, the procedure go through the $ log(len)+1 $ layers of STree, in each layer it updates the snapshot of node covering $ t $ or creates necessary node first. Since updating a snapshot and creating a new empty node both contain a constant number of operations, the time complexity of \kw{STreeInsertEdge} procedure is $ O(log(len)+1) $.

\section{Effectiveness of Indexes} \label{sec:effec_of_indexes}
In common range query problems, indexes are helpful to accelerate the queries. However, similar indexes can not always work in our problem, since the elements in the array are not numeric values but snapshots.

Consider the first example in Figure~\ref{fig:index_fail_success_example} (a),
\begin{figure}[t] %\vspace*{-0.5cm}
	\begin{center}
		\begin{tabular}[t]{c}\hspace*{-0.3cm}
			\subfigure[{\scriptsize }]{
				\includegraphics[width=0.95\columnwidth]{fig/example/example_index_fail.pdf}
			}
			\hspace*{-0.3cm}
			\\
			\subfigure[{\scriptsize }]{
				\includegraphics[width=0.95\columnwidth]{fig/example/example_index_success.pdf}
			}
		\end{tabular}

		\caption{Two cases of range query on snapshot array}
		\vspace*{-0.4cm}
		\label{fig:index_fail_success_example}
	\end{center}
\end{figure}
there is no single edge shared by any two snapshots (\textbf{edge $ e $ is shared by snapshots means there exists an edge that share the vertices with $ e $ in each of those snapshots}). In this case, if we merge snapshots in $ [t_1,t_4] $, the best choice is the baseline method, i.e., traversing all snapshots in the range and computing the result one by one. In baseline method, only $ 4 $ edges will be visited and each of them is visited only once. The edges in the result (snapshot at the right of the arrow in Figure~\ref{fig:index_fail_success_example} (a)) are exactly those $ 4 $ edges, so baseline method achieves the highest efficiency and there is no need to build indexes in this case.

Consider the second example in Figure~\ref{fig:index_fail_success_example} (b), all snapshots share the same edges. In this case, if we still merge snapshots in $ [t_1,t_4] $, each edge will be visited $ 4 $ times in baseline method, so it's necessary to build indexes. Next, we discribe the measurement of the necessity of indexes and the key factor we found which affects the necessity of indexes.
\begin{definition}[Acceleration Ratio of Index]
	\label{def:acceleration_ratio_index}
	For an array $ Arr $ holding numeric values or snapshots, suppose we build one kind of index on it. For a query $ q=(func,[l,r]) $, suppose the basic components are visited $ C_b $ times using baseline method and $ C_i $ times using index, the acceleration ratio of the index on $ q $ is $ \frac{C_b}{C_i} $.
\end{definition}
The basic components in snapshot array and indexes built on snapshot array are edges. We use the number of visited edges as the time consumption of conducting a query using baseline method and using indexes in Definition~\ref{def:acceleration_ratio_index}, because from previous analyses we can find that visiting edges is the main part of the execution of query algorithms of baseline method and indexes.
\begin{definition}[Similarity of Snapshots]
	\label{def:similarity_snapshots}
	Given a snapshot array $ snapArr $ starting from $ 1 $ and containing $ n $ snapshots, the summative snapshot $ S $ of all snapshots in $ snapArr $, $ E=\sum_{i=1}^{n}|snapArr[i]| $, the similarity of snapshots in $ snapArr $ is $ s=\frac{E}{|S|} $.
\end{definition}
In other words, the similarity of snapshots is the ratio of the total number of edges in snapshot array to the size of the summative snapshot. For simplicity, we assume that each snapshot in $ snapArr $ has at least one edge. Clearly, for a snapshot array of length $ n $, similarity of snapshots $ s $ is in $ [1,n] $.
\begin{example}
	Consider examples in Figure~\ref{fig:similarity}.
	\begin{figure}[t] %\vspace*{-0.5cm}
		\begin{center}
			\begin{tabular}[t]{c}\hspace*{-0.3cm}
				\subfigure[{\scriptsize $ s=1 $}]{
					\includegraphics[width=0.5\columnwidth]{fig/example/example_similarity_a.pdf}
				}
				\hspace*{-0.5cm}
				\subfigure[{\scriptsize $ s=4 $}]{
					\includegraphics[width=0.5\columnwidth]{fig/example/example_similarity_b.pdf}
				}
				\\
				\subfigure[{\scriptsize $ s=\frac{5}{3} $}]{
					\includegraphics[width=0.5\columnwidth]{fig/example/example_similarity_c.pdf}
				}
				\hspace*{-0.5cm}
				\subfigure[{\scriptsize $ s=4 $}]{
					\includegraphics[width=0.5\columnwidth]{fig/example/example_similarity_d.pdf}
				}
			\end{tabular}
			\vspace*{-0.3cm}
			\caption{Four snapshot arrays with different similarity of snapshots}
			\vspace*{-0.4cm}
			\label{fig:similarity}
		\end{center}
	\end{figure}
	There are 4 snapshot arrays having the same length of 4 but different similarities of snapshots $ s $. In Figure~\ref{fig:similarity} (a), $ s=1 $, and we can see that there is no edge shared by any two snapshots, or, snapshots in Figure~\ref{fig:similarity} (a) are not similar to each other. In Figure~\ref{fig:similarity} (b) and (d), $ s=4 $, all snapshots share the same edges. In Figure~\ref{fig:similarity} (c), $ s=\frac{5}{3} $, and there are edges shared by some snapshots.
\end{example}
From examples above we can see that with the similarity of snapshots grows, there will be more snapshots in snapshot array which are similar to each other, or, more edges are shared by snapshots.

\begin{theorem}
	\label{thm:best_index_efficiency}
	For two arrays of length $ n $ and starting from $ 1 $, snapshot array $ snapArr $ and numeric array $ numArr $, if similarity of snapshots in $ snapArr $ equals $ n $, then for any query $ q=(func,[l,r]) $ on both $ snapArr $ and $ numArr $, the acceleration ratios of the same index built on them are equal.
\end{theorem}
\begin{proof}
	In Section~\ref{sec:indexes}, we know that the elements in all indexes built on snapshot array are snapshots, which can be regarded as summative snapshots of different snapshot sub-arrays of the original snapshot array. If the similarity of snapshots in $ snapArr $ is $ n $, then all snapshots in $ snapArr $, as well as their summative snapshot, share the same edges. It's clear that all snapshots in indexes also share the same edges with snapshots in $ snapArr $. Suppose the indexes built on $ snapArr $ and $ numArr $ are two STrees $ snapSTree $ and $ numSTree $ (the proofs on prefix array and STable are simple, so we only analyse STree here), and summative snapshot of snapshots in $ snapArr $ is $ S $. In baseline method, conducting $ q $ requires $ r-l+1 $ visits on numeric values in $ numArr $ and $ (r-l+1)|S| $ visits on edges in $ snapArr $. In STree, the structures of $ snapSTree $ and $ numSTree $ are the same, because the structure of STree is only related to the entire range of the array maintained by STree, and nodes where merging happens in $ snapSTree $ and $ numSTree $ (when condition in line~2 of Algorithm~\ref{alg:stree_query} is true) correspond one by one, because the ways that $ [l,r] $ in $ q $ split are the same in $ snapSTree $ and $ numSTree $. If there are $ C_n $ nodes being merged in $ numSTree $, then there are $ C_n $ nodes being merged in $ snapSTree $ and $ C_n|S| $ edges being visited. Thus, the acceleration ratio of $ snapSTree $ ($ \frac{(r-l+1)|S|}{C_n|S|} $) is equal to that of $ numSTree $ ($ \frac{r-l+1}{C_n} $).
\end{proof}
\begin{theorem}
	\label{thm:worst_index_efficiency}
	For a snapshot array $ snapArr $ of length $ n $ and starting from 1. If similarity of snapshots in $ snapArr $ equals 1, then for query $ q=(func,[l,r]) $ on $ snapArr $, the acceleration ratios of STable and STree are less than or equal to 1, and the acceleration ratio of prefix array is $ \frac{\sum_{k=l}^{r}|snapArr[k]|}{\sum_{k=1}^{r}|snapArr[k]|} $.
\end{theorem}
\begin{proof}
	Suppose $ l,r $ are biased timestamps. Since the similarity of snapshots in $ snapArr $ is 1, there is no edge shared by any snapshots in $ snapArr $, e.g., snapshot array in Figure~\ref{fig:index_fail_success_example} (a). For any summative snapshot $ S_{ij} $ of sub-array $ [snapArr[i],snapArr[i+1],...,snapArr[j]] $, we can easily get $ |S_{ij}|=\sum_{k=i}^{j}|snapArr[k]| $. For STree $ snapSTree $ built for $ snapArr $ (time range in $ q $ should be $ [l+t_{base},r+t_{base}] $ when using STree, and the proof on STable is simple so we omit it here), suppose the set of nodes being merged in conducting $ q $ is $ NS $, we can easily get $ \sum_{node \in NS}|node.snapshot|=\sum_{k=l}^{r}|snapArr[k]| $, where $ \forall node \in NS,|node.snapshot|=\sum_{k=node.timel}^{node.timer}|snapArr[k-t_{base}]| $. As a result, in query functions of baseline method and STree, we must visit the same number of edges. Considering the cost of visiting nodes in STree, the acceleration ratio of STree is less than or equal to 1. For prefix array $ preArray $ built for $ snapArr $, as in Algorithm~\ref{alg:prefix_array_query}, the number of edges being visited in conducting $ q $ is $ |preArray[r]|=\sum_{k=1}^{r}|snapArr[k]| $. Thus, the acceleration ratio of prefix array is $ \frac{\sum_{k=l}^{r}|snapArr[k]|}{\sum_{k=1}^{r}|snapArr[k]|} $.
\end{proof}

The above two theorems make clear the necessity of indexes in two extreme cases. As in Theorem~\ref{thm:best_index_efficiency}, if similarity of snapshots in snapshot array equals the length of the snapshot array, all indexes achieve the designed acceleration ratio as they perform in range query on numeric arrays. As in Theorem~\ref{thm:worst_index_efficiency}, if similarity of snapshots in snapshot array equals 1, STable and STree have no acceleration effect and prefix array may even have negative effect. Baseline method is the best way to conduct any query in this case, because the number of edges in the result is exactly the number of edges being visited in baseline method.

It's difficult to model the relationship between similarity of snapshots and acceleration ratio of each index on all possible queries precisely, because similarity of snapshots discribes the overall similarity of snapshots in snapshot array, but in specific queries, snapshots in sub-arrays specified by ranges of the queries may not show the same similarity as the overall similarity. However, it's possible to model the relationship roughly. If the similarity of snapshots in snapshot array equals 1, then there is no edge shared by two snapshots and visited more than once in baseline method. If the similarity of snapshots in snapshot array equals the length of the array, then all edges are shared by all snapshots. In this case, each edge will be visited repeatedly in each cycle of baseline method (line~8 in Algorithm~\ref{alg:temp_cuboid_baseline}), while indexes store summative snapshots of some sub-arrays of the snapshot array in advance to reduce the number of repeated visits on edges. With the similarity of snapshots grows from 1 to the length of snapshot array, more edges will be shared by snapshots, making the efficiency of baseline method lower and efficiency of indexes higher. As a result, the acceleration ratio of indexes grows when similarity of snapshots grows, which will be proved in experiments in Section~\ref{subsec:efficiency_eva}.

The similarity of snapshots also affects the space consumption of indexes. Since the elements in all snapshot arrays and all indexes built on them are snapshots, we use the total number of edges in all snapshots to indicate the space consumption. For example, if we build two STrees $ STree_a $ and $ STree_b $ on both snapshot arrays in Figure~\ref{fig:similarity} (a) and (b), the snapshot in each node of $ STree_b $ has only one edge, while the snapshot in each node of $ STree_a $ may have to store more than one edge, even if Figure~\ref{fig:similarity} (a) and (b) have the same number of edges.

\begin{definition}[Space Consumption Ratio]
	\label{def:space_consum_ratio}
	For an array $ Arr $ holding numeric values or snapshots, suppose we build one kind of index on it. If $ Arr $ is a snapshot array, let the total number of edges in $ Arr $ is $ E_a $, and the total number of edges in the index is $ E_i $. If $ Arr $ is a numeric array, $ E_a,E_i $ are the total number of numeric values in $ Arr $ and in the index respectively. The space consumption ratio of the index is $ \frac{E_i}{E_a} $.
\end{definition}
\begin{theorem}\label{thm:best_index_space_ratio}
	For two arrays of length $ n $ and starting from 1, snapshot array $ snapArr $ and numeric array $ numArr $, if similarity of snapshots in $ snapArr $ equals $ n $, then the space consumption ratios of the same index built on the two arrays are equal.
\end{theorem}
\begin{proof}
	We mentioned that in proof of Theorem~\ref{thm:best_index_efficiency}, the structures of STree built on $ numArr $ and $ snapArr $ are the same. It is the same with the array structure of prefix array and STable. We can turn indexes into sets holding elements like: $ (SumElement,[l,r]) $, and if an index is built on $ numArr $, $ SumElement $ is the aggregation result of numeric values in $ [l,r] $, otherwise the summative snapshot of snapshots in $ [l,r] $. For each index, suppose we build the index on $ numArr $ and $ snapArr $, named $ numIndex $ and $ snapIndex $. There must be a bijection between elements in $ numIndex $ and $ snapIndex $ that for each $ i $ in $ numIndex $, there is a corresponding $ j $ in $ snapIndex $ that $ i $ and $ j $ have the same time range $ [l,r] $, since the structures of $ numIndex $ and $ snapIndex $ are the same. It's easy to know that the space consumption ratio of $ numIndex $ is $ \frac{|numIndex|}{n} $. Since the similarity of snapshots in $ snapArr $ is $ n $, all snapshots in $ snapArr $ and $ snapIndex $ share the same edges. Suppose the summative snapshot of snapshots in $ snapArr $ is $ S $, the space consumption ratio of $ snapIndex $ is $ \frac{|snapIndex||S|}{n|S|} $, since $ |snapIndex|=|numIndex| $, so the space consumption ratio of $ snapIndex $ and $ numIndex $ are equal.
\end{proof}

Similar to acceleration ratio of indexes, the space consumption ratio of indexes is affected negatively by the decrease of similarity of snapshots. Specifically we discuss the space consumption ratio of STree if similarity of snapshots equals 1.
\begin{theorem}\label{thm:stree_worst_index_space_ratio}
	For a snapshot array $ snapArr $ of length $ n $ and starting from 1. If similarity of snapshots in $ snapArr $ equals 1, then the space consumption ratio of STree built on $ snapArr $ is in $ (\lceil log(n) \rceil,\lceil log(n) \rceil+1] $.
\end{theorem}
\begin{proof}
	Let $ m=\sum_{i=1}^{n}|snapArr[i]| $. If $ log(n) $ is an integer, the height of STree built on $ snapArr $ is $ log(n)+1 $. Consider the $ (log(n)+1) $-th layer of STree, all nodes on the $ (log(n)+1) $-th layer of STree are leaves. It's clear that the total number of edges in snapshots of the leaves is $ m $. For each node on the $ log(n) $-th layer, the snapshot in the node is the summative snapshot of snapshots in the two siblings of the node. Since the similarity of snapshots in $ snapArr $ is 1, no edge is shared by snapshots in $ snapArr $, so the size of snapshot of each node on the $ log(n) $-th layer is the sum of the size of snapshots in the two siblings. Thus, the total number of edges in snapshots of nodes on the $ log(n) $-th layer is still $ m $. It's the same for nodes on each layer. Finally, the total number of edges in snapshots of all nodes in STree is $ (log(n)+1)\times m $, so the space consumption ratio of STree is $ log(n)+1 $.

	If $ log(n) $ is not an integer, then the number of leaves on the $(\lceil log(n) \rceil +1)$-th layer is less than $ n $, so there are less than $ m $ edges on the $(\lceil log(n) \rceil +1)$-th layer, but still $ m $ edges on other layers. Thus, the total number of edges in STree is in $ (\lceil log(n) \rceil\times m, (\lceil log(n) \rceil+1)\times m) $. As a result, in this case, the space consumption ratio of STree is in $ (\lceil log(n) \rceil , \lceil log(n) \rceil+1) $.
\end{proof}

From Theorem~\ref{thm:stree_worst_index_space_ratio}, if similarity of snapshots equals 1, we need more space to store STree, which may be $ log(n)+1 $ times the space of the original snapshot array, while only 2 times the space is needed to store STree if similarity of snapshots equals the length of the original snapshot array. For prefix array and STable, it's impossible to derive a clear mathematical expression of the space consumption ratio related to $ n $ only. But similar to STree, with the decrease of similarity of snapshots, each snapshot in indexes has to store more edges even if the number of edges in its corresponding sub-array of the original snapshot array remains the same. We will prove the above findings in experiment of Section~\ref{subsec:efficiency_eva}.

From analysis above, we can conclude that the necessity of building indexes, which can be determined by acceleration ratio and space consumption ratio of indexes, varies with the similarity of snapshots. This is the biggest difference of range query on snapshot array compared to range query on numeric array. Based on this observation, we'd better build indexes on snapshot array with high similarity of snapshots, which is compatible with the implementation strategy of \kw{Temporal\enspace Graph\enspace Cube} presented in the next section, which requires precomputing views with high similarity of snapshots in their snapshot arrays.
%In particular, we use, we know that if similarity of snapshots $ s $ equals $ 1 $, baseline method is the best method and there is no need to build indexes. If $ s $ equals the length of the snapshot array, indexes achieve the designed acceleration ratio, so it's necessary to build indexes just like what we usually do in common range query problem. However, in real-world temp-multi-networks and their precomputed temp-aggre-networks, similarities of snapshots in snapshot arrays are hardly to be exactly $ n $ or $ 1 $ ($ n $ is the length of snapshot arrays). For a fixed length $ n $, with $ s $ drops from $ n $ to $ 1 $, it's more likely that the number of reduplicate edges visited in baseline method compared to the size of result snapshot decrease in any query, making baseline method more efficient and indexes more unnecessary.

%It's difficult to model the relationship between similarity of snapshots and acceleration ratio of each index on all possible queries precisely, because similarity of snapshots discribes the overall similarity of snapshots in snapshot array, but in specific queries, snapshots in ranges of the queries may not show the same similarity as the overall similarity. So in Section~\ref{sec:experiment}, we mainly discuss the relationship between similarity of snapshots and the overall acceleration ratio of each index on all possible queries.[deffinition of overall?]

%\vspace{-0.2cm}
\section{Partial Materialization} \label{sec:partial_materialization}
To implement a temporal graph cube, we need precompute, or materialize some or all views in the temporal graph cube, since precomputed views can reduce the response time of OLAP queries and operations like roll-up, drill-down and slice-and-dice as we discussed in Section~\ref{subsec:queries}.

Like static graph cube, there are three possible alternatives: 1) Full materialization. 2) No materialization. 3) Partial materialization. The first and the second alternatives are both unrealistic, because: 1) Full materialization means we need to precompute all possible views, but the number of views grows exponentially with $ |dim(A_{base})| $. It's work only if $ |dim(A_{base})| $ is small enough. 2) Not precomputing any view (except $ A_{base} $) leads to long response time in almost all OLAP queries and operations.

For large temp-multi-networks, we can only choose partial materialization, i.e., select a set of views to be precomputed in order to balance the space consumption and average response time according to the probability distribution of all possible queries. \textit{View selection} in traditional data cube scenario is a NP-hard problem \cite{karloff1999complexity}. It is also a NP-hard problem in static graph cube scenario because traditional data cube can be ragarded as a special case of static graph cube \cite{zhao2011graph}. Similarly, static graph cube can also be ragarded as a special case of temporal graph cube, if only one timestamp exists in the temporal graph cube. Thus, \textit{view selection} problem in temporal graph cube is also a NP-hard problem. In traditional data cube and static graph cube, \textit{view selection} problem is usually solved by heuristic sub-optimal solutions \cite{karloff1999complexity,zhao2011graph}, such as greedy algorithm and its variations \cite{morfonios2007rolap}. Those heuristic solutions always measure the effectiveness of the selected views by benefit those views bring.

\textit{Benefit} was first introduced in \cite{harinarayan1996implementing}, we briefly review the definition of \textit{benefit} below.
\begin{definition}[Benefit of Selected Views]
	\label{def:benefit_views}
	For a sequence of selected views $ u_1,u_2,...,u_{k-1} $ in the order of being selected ($ u_1 $ is always $ A_{base} $), the benefit of the candidate view $ u $ to be selected next is $ B(u,S_{k-1}) $, i.e., benefit brought by $ u $ w.r.t. $ S_{k-1}=\{u_i|i \in [1,k-1]\}$. $ B(u,S_{k-1}) $ is defined as follows:
	\begin{enumerate}
		\item For each $ w \preceq u $, define the quantity $ B_w $ by:
		\begin{enumerate}[label={(\alph*)}]
			\item Let $ v $ be the view of least cost in $ S_{k-1} $ and $ w \preceq v $.
			\item If $ C(u)<C(v) $, then $ B_w=C(v)-C(u) $. Otherwise $ B_w=0 $.
		\end{enumerate}
		\item $ B(u,S_{k-1})=\sum_{w \preceq u}B_w $.
	\end{enumerate}
	$ C(\ast) $ returns the cost of views. In traditional data cube and static graph cube scenario, cost of views is the size of their corresponding precomputed tables or networks. The total benefit of selected view sequence is $ \sum_{i=2}^{k}B(u_i,S_{i-1}) $.
\end{definition}
In greedy algorithm, among all candidate views, the view $ u $ with the biggest $ B(u,S_{k-1}) $ is selected to be $ u_k $. For the total benefit of view squence generated by each heuristic solution, the closer to the benefit of the optimal view sequence, the more effective the solution is.

In Definition~\ref{def:benefit_views}, the cost of views is used in representation of the benefit, because the cost of views is also the time complexity or cost of queries in data cube and static graph cube. However, cost of views in temporal graph cube (space consumption of the corresponding temporal aggregate networks of views) can not be used as the cost of temporal cuboid queries. First, in Definition~\ref{def:benefit_views}, $ C(u) $ can not be the cost of temporal cuboid query $ Q=(w,[l,r]) $ if we conduct $ Q $ on $ u $. It's similar to $ C(v) $. The reason is that there is a specified time range $ [l,r] $ in $ Q $, which doesn't exist in any query of traditional data cube or static graph cube. In implementation of temporal graph cube, we know nothing about the specified time ranges in all possible queries. Second, if we use an expected range $ [l_e,r_e] $ as the representative of all ranges in all queries according to the probability distribution of specified ranges in all queries, we may be able to get the time complexity of conducting $ Q=(w,[l_e,r_e]) $ on $ u $ using baseline method by precomputing the total size of snapshots in $ [l_e,r_e] $ and the size of summative snapshot of snapshots in $ [l_e,r_e] $. Only in this way can we get a relatively accurate benefit for each $ u $. Third, the above precomputing process is very costly, and it is based on conducting all queries with baseline method, but in Section~\ref{sec:effec_of_indexes} and \ref{sec:experiment}, we prove the effectiveness of indexes in specific cases, so the cost of temporal cuboid queries is determined by the indexes. However, it's more costly to precompute the time complexity of conducting $ Q=(w,[l_e,r_e]) $ on $ u $ using indexes of $ u $, not to mention that timestamps will be added into temporal networks constantly, making that we can not find a stable $ [l_e,r_e] $.

Thus, greedy algorithm and its variations are not suitable for view selection problem in temporal graph cube. In this paper, we adopt a much simpler solution to select views to be precomputed, \kw{MinLevel}, which is introduced in \cite{zhao2011graph}. The idea of \kw{MinLevel} is simple: users are more willing to query with small number of dimensions, i.e., $ |dim(A)| $ is small in $ Q=(A,[l,r]) $. In \kw{MinLevel}, view $ A $ where $ |dim(A)|=l_0 $ is in the first batch of views to be selected ($ l_0 $ is an empirical value). If all the views with $ l_0 $ dimensions have been selected and the number of selected views or the total size needed for precomputing the selected views hasn't reach the limit, then we continue to select views with $ l_0+1 $ dimensions util we select enough number or size of views. In this paper we choose to select $ k $ views to be precomputed and $ l_0=1 $.

Using \kw{MinLevel}, we precompute views with small number of dimensions, and these views also have high similarity of snapshots in their snapshot arrays (as shown in Table~\ref{tab:materialized_view}), which is a good news for indexes as mentioned in Section~\ref{sec:effec_of_indexes}. In conducting temporal cuboid query $ Q=(A',[l,r]) $ in Section~\ref{subsec:queries}, if $ A' $ is not precomputed and $ l_0=1 $, only $ A_{base} $ can be used to conduct $ Q $. If $ l_0>1 $ and $ |dim(A')|<l_0 $, then there may be some precomputed views $ \{A_1'',A_2'',...\} $ where $ |dim(A_i'')|=l_0 $ can be used to conduct $ Q $. In this case, choosing an arbitrary $ A_i'' $ to conduct $ Q $ is efficient enough because the size of $ A_i'' $ and its indexes are usually small since $ l_0 $ is a small value.

\section{Experiment} \label{sec:experiment}
In this section, we conduct extensive experiments to evaluate the effectiveness and efficiency of the temporal graph cube model.

\subsection{Datasets} \label{subsec:datasets}
We conduct experiments on two real-world temp-multi-networks: DBLP (\url{https://dblp.org/xml}) and IMDB (\url{https://www.imdb.com/interfaces}), to evaluate the effectiveness and efficiency of the temporal graph cube respectively. Below we introduce the details of the two datasets.

\begin{table}[t!]%\vspace*{-0.3cm}
	%\scriptsize
	\small
	\centering
	\caption{Basic information of datasets} \vspace*{-0.2cm} \label{tab:datasets}
	\begin{tabular}{c c c c c}
		\toprule
		\bf Dataset & $|\Vr|$ & $|\Er|$ & $ t_{end} $ & Time scale \cr %\hline %\hline
		\midrule
		\kw{DBLP} & 1,242,883 & 9,532,533 & 85 & year \cr %\hline
		\kw{IMDB} & 3,497,300 & 39,682,231 & 151 & year  \cr %\hline
		\bottomrule
	\end{tabular}
	\vspace*{-0.2cm}
\end{table}

\stitle{DBLP.} This dataset is extracted from dblp computer science bibliography data which is downloaded in September, 2021. Specifically, we extract publication information of all conferences and journals listed in \textit{Catalogue of international academic conferences and periodicals recommended by China Computer Federation} (\url{https://www.ccf.org.cn/ccf/contentcore/resource/download?ID=99185}) except conferences and journals which focus on cross field since we can not classify them into AI, DB or other specific fields. We gain $ 545 $ conferences and journals in total, covering $ 9 $ fields in computer science: CA(computer architecture), CN(computer network), NS(network security), SE(software engineering), DB(database), TH(computer science theory), CG(computer graphics), AI(artificial intelligence) and HI(human-computer interaction). Table~\ref{tab:datasets} illuminates the basic information of temp-multi-network based on DBLP dataset. For each vertex(author), there are three dimensions of information: \kw{Name}, \kw{Area}, \kw{Productivity}. Since we classify all publications into $ 9 $ areas, \kw{Area} of an author is one of the $ 9 $ areas where he or she has the most publications ever, which can be regarded as the representative of focus of his or her works. For \kw{Productivity}, we discretize the number of publications per year of an author into four different buckets, which is shown in Table~\ref{tab:productivity}. The number of publications per year of an author equals the total number of publications of the author divided by $ 2022-y $, $ y $ is the first year when the author had his or her first publication. \kw{Productivity} specifies the ability of producing academic papers of an author.
\begin{table}[t!]%\vspace*{-0.3cm}
	%\scriptsize
	\small
	\centering
	\caption{Discretization of publication number per year} \vspace*{-0.2cm} \label{tab:productivity}
	\begin{tabular}{c c}
		\toprule
		\bf Productivity & Publication number per year $ x $ \cr %\hline %\hline
		\midrule
		Excellent & $ 8<x $ \cr %\hline
		Good & $ 3 < x \leq 8 $  \cr %\hline
		Fair & $ 1 < x \leq 3 $  \cr %\hline
		Poor & $ x \leq 1 $  \cr %\hline
		\bottomrule
	\end{tabular}
	\vspace*{-0.4cm}
\end{table}
Each temporal edge represents that two authors coauthored a paper at a certain time (year).

%Note that there is no numeric attribute in dimensions of vertices of temp-multi-network in \kw{DBLP}, so we can only specify $ f_v= $ COUNT($\ast$). It's not important because the processes of aggregating on vertices are the same when $ f_v $ varies. Temporal edges in \kw{DBLP} also have no numeric attribute, so $ f_e $ can only be COUNT($\ast$). Fortunately, there are numeric attributes in temporal edges of the second dataset.

\stitle{IMDB.} This dataset is extracted from the Internet Movie Database (IMDB) downloaded in November, 2021. The IMDB dataset contains various types of works like movie, short, video, etc. Each work has the following attributes: \kw{tile}, \kw{time}, \kw{titleType}, \kw{genre}, \kw{isAdult}, \kw{rating}, \kw{voteNum}. \kw{rating} is a numeric attribute, representing the average rating for the work from individual users. \kw{voteNum} is the number of votes the work has received. There are some other files about principals of each work, which will be used as vertices of the temporal network. In summary, we extract several dimensions for each principal: \kw{Name}, \kw{Pro}, \kw{Field}, \kw{Genre}, \kw{isAdult}, \kw{Averating}, \kw{Hotness}. \kw{Pro} is the primary profession of a person, e.g., director, actor, producer, soundtrack, etc., which can be directly extracted from certain files. \kw{Field} specifies the field that a person belongs to, e.g., movie, short, video, etc., which is obtained by counting the types of the works that the person participates in and choosing the type having the biggest count. \kw{Genre} specifies the genre of works that a person usually participates in, which is obtained by the same process in obtaining \kw{Field}. \kw{isAdult} specifies if a person mainly works for adult works or not. \kw{Averating} specifies the average rating of works that a person participates in, and it shows the average quality of works related to the person. \kw{Hotness} specifies the hotness of works a person usually participates in, which is expressed by the average vote number of works related to the person. \kw{Averating} and \kw{Hotness} both need to be discretized, which are shown in Table~\ref{tab:averating} and Table~\ref{tab:hotness} respectively.
\begin{table}[t!]%\vspace*{-0.3cm}
	%\scriptsize
	\small
	\centering
	\caption{Discretization of average rating} \vspace*{-0.2cm} \label{tab:averating}
	\begin{tabular}{c c}
		\toprule
		\bf Averating & \kw{Averating} value $ x $ \cr %\hline %\hline
		\midrule
		5 & $ 8 \leq x $ \cr %\hline
		4 & $ 6 \leq x < 8 $  \cr %\hline
		3 & $ 4 \leq x < 6 $  \cr %\hline
		2 & $ 2 \leq x < 4 $  \cr %\hline
		1 & $ x < 2 $  \cr %\hline
		\bottomrule
	\end{tabular}
	\vspace*{-0.3cm}
\end{table}
\begin{table}[t!]%\vspace*{-0.3cm}
	%\scriptsize
	\small
	\centering
	\caption{Discretization of average vote number} \vspace*{-0.2cm} \label{tab:hotness}
	\begin{tabular}{c c}
		\toprule
		\bf Hotness & Average vote number of works $ x $ \cr %\hline %\hline
		\midrule
		hot & $ 100,000 \leq x $ \cr %\hline
		popular & $ 10,000 \leq x < 100,000 $  \cr %\hline
		fair & $ 1,000 \leq x < 10,000 $  \cr %\hline
		cold & $ x < 1,000 $  \cr %\hline
		\bottomrule
	\end{tabular}
	\vspace*{-0.4cm}
\end{table}

We use persons (principals in works) as the vertices in the temp-multi-network, and two vertices can form a temporal edge if they work together for a same work. The timestamp of the temporal edge is the release time of the work, and the numeric attribute of the temporal edge is the average rating of the work. The statistical information of temp-multi-network in IMDB is shown in Table~\ref{tab:datasets}.

\subsection{Effectiveness Evaluation} \label{subsec:effectiveness_eva}
In this section we evaluate the effectiveness of temporal graph cube as a decision-support tool on \kw{DBLP}. We will conduct some queries and show some interesting findings, especially time-dependent findings. First, we conduct a series of temporal cuboid queries $ Q_1=((\kw{Area}),[2001,2006]) $, $ Q_2=((\kw{Area}),[2006,2011]) $, $ Q_3=((\kw{Area}),[2011,2016]) $, $ Q_4=((\kw{Area}),[2016,2021]) $ on \kw{DBLP}. Aggregation functions upon vertices and edges are both COUNT($\ast$).
\begin{table}[t!]%\vspace*{-0.3cm}
	%\scriptsize
	\small
	\centering
	\caption{Vertex table of $ Q_1-Q_4 $} \vspace*{-0.2cm} \label{tab:vertex_table_dblp_area}
	\begin{tabular}{c c}
		\toprule
		Area & COUNT($\ast$) \cr %\hline %\hline
		\midrule
		CA & 228,429 \cr %\hline
		CN & 156,329  \cr %\hline
		NS & 60,880  \cr %\hline
		SE & 114,727  \cr %\hline
		DB & 102,931  \cr %\hline
		TH & 109,293  \cr %\hline
		CG & 156,982  \cr %\hline
		AI & 256,048 \cr %\hline
		HI & 57,264  \cr %\hline
		\bottomrule
	\end{tabular}
	\vspace*{-0.4cm}
\end{table}
\begin{figure*}[t!]
	\begin{center}
		\begin{tabular}[t]{c}\hspace*{-0.3cm}
			\subfigure[{\scriptsize $ Q_1 $}]{
				\includegraphics[width=0.23\textwidth]{fig/exp/visio_area_2001_2006.pdf}
			}
			\hspace*{-0.3cm}
			\subfigure[{\scriptsize $ Q_2 $}]{
				\includegraphics[width=0.23\textwidth]{fig/exp/visio_area_2006_2011.pdf}
			}
			\hspace*{-0.3cm}
			%\vspace*{-0.3cm}\\\hspace*{-0.3cm}
			\subfigure[{\scriptsize $ Q_3 $}]{
				\includegraphics[width=0.23\textwidth]{fig/exp/visio_area_2011_2016.pdf}
			}
			\hspace*{-0.3cm}
			\subfigure[{\scriptsize $ Q_4 $}]{
				\includegraphics[width=0.23\textwidth]{fig/exp/visio_area_2016_2021.pdf}
			}
		\end{tabular}
	\end{center}
	\vspace*{-0.4cm}
	\caption{Network structures in the results of queries $ Q_1-Q_4 $}
	\vspace*{-0.4cm}
	\label{fig:dblp_area}
\end{figure*}
The results are shown in Table~\ref{tab:vertex_table_dblp_area} and Figure~\ref{fig:dblp_area}. Note that on time scale of years, an author may just be born or pass away at some time. But such information is not contained in any dataset we know. Also, we have no way to know when a researcher starts and ends his or her research career, so in this experiment, the COUNT($\ast$) values in vertex table shared by $ Q_1 $-$ Q_4 $ are not very informative, and we will focus on the network structures of $ Q_1 $-$ Q_4 $ and numeric information on them. The network structures of $ Q_1 $-$ Q_4 $ are shown in Figure~\ref{fig:dblp_area}, illustrating the co-authorship patterns between authors grouped by different areas in the specified time ranges. For simplicity, in Figure~\ref{fig:dblp_area} (a) we omit those co-authorships with COUNT($\ast$) values less than 10,000, which are also omitted in Figure~\ref{fig:dblp_area} (b)-(d), regardless of their COUNT($\ast$) values.

Some interesting phenomenons can be found in the results. For example, authors in each area co-author most with authors in the same area even in different time ranges. As for co-authorship between authors in different areas, for example, we can easily find that in recent 20 years, the number of co-authorship between each pair of areas increases over time. We can also find in recent 20 years researchers in \kw{AI} co-authored with researchers in \kw{CG} most. This is because there are many problems in computer graphics like automatic modeling, image processing and digital photography which can be solved by AI technology. More interestingly, in $ [2001,2006] $, researchers in \kw{TH} co-authored most with researchers in \kw{CA}, but with time passes, researchers in \kw{TH} are more willing to co-author with researchers in \kw{AI}. The reason may be that theoretical research on Artificial Intelligence becomes more and more important with the wide application of Artificial Intelligence in recent years. The above findings can also be obtained from uncut results.
%[ns and cn,ca and ai;ai and cg;th and ai,ca]

In each peroid we can zoom into a more fine-grained view by conducting a drill-down operation. For example, for $ Q_4=((\kw{Area}),[2016,2021]) $, drill-down can be a query $ Q_4'=((\kw{Area},\kw{Productivity}),[2016,2021]) $. The network structure of $ Q_4' $ is shown in Figure~\ref{fig:dblp_area_pro_2016_2021}.
\begin{figure}
	\begin{center}
		\includegraphics[width=0.95\columnwidth]{fig/exp/visio_area_pro_2016_2021.pdf}
	\end{center}

	\vspace*{-0.4cm}
	\caption{Network structure in the result of $ Q_4' $}
	\vspace*{-0.4cm}
	\label{fig:dblp_area_pro_2016_2021}
\end{figure}
For simplicity, we drop edges with weight less than 10,000 and we only keep vertices related to areas in $ \{\kw{AI},\kw{DB}\} $ (it can be regarded as a slice-and-dice operation). In Figure~\ref{fig:dblp_area} (d), there are 117,338 co-authorships between reseachers of \kw{AI} and \kw{DB} in $ [2016,2021] $. In fine-grained view of Figure~\ref{fig:dblp_area_pro_2016_2021}, we can see that the most co-authorships belong to researchers of poor and fair productivity in \kw{AI} and \kw{DB}.

We then examine the effectiveness of temporal crossboid query.
\begin{figure*}[t!]
	\begin{center}
		\begin{tabular}[t]{c}\hspace*{-0.3cm}
			\subfigure[{\scriptsize $ Q_{cross1} $}]{
				\includegraphics[width=0.25\textwidth]{fig/exp/dblp_cross_name_area_2001_2006.pdf}
			}
			\hspace*{-0.3cm}
			\subfigure[{\scriptsize $ Q_{cross2} $}]{
				\includegraphics[width=0.25\textwidth]{fig/exp/dblp_cross_name_area_2006_2011.pdf}
			}
			\hspace*{-0.3cm}
			%\vspace*{-0.3cm}\\\hspace*{-0.3cm}
			\subfigure[{\scriptsize $ Q_{cross3} $}]{
				\includegraphics[width=0.25\textwidth]{fig/exp/dblp_cross_name_area_2011_2016.pdf}
			}
			\hspace*{-0.3cm}
			\subfigure[{\scriptsize $ Q_{cross4} $}]{
				\includegraphics[width=0.25\textwidth]{fig/exp/dblp_cross_name_area_2016_2021.pdf}
			}
		\end{tabular}
	\end{center}
	\vspace*{-0.4cm}
	\caption{Network structures in the results of $ Q_{cross1}-Q_{cross4} $}
	\vspace*{-0.4cm}
	\label{fig:dblp_cross}
\end{figure*}
Figure~\ref{fig:dblp_cross} shows the results of a serise of temporal crossboid queries:
\begin{center}
	$ Q_{cross1}=((\kw{Name}),(\kw{Area}),[2001,2006]) $,\\$ Q_{cross2}=((\kw{Name}),(\kw{Area}),[2006,2011]) $,\\$ Q_{cross3}=((\kw{Name}),(\kw{Area}),[2011,2016]) $,\\$ Q_{cross4}=((\kw{Name}),(\kw{Area}),[2016,2021]) $.
\end{center}
The above queries all focus on the cross interaction between \kw{Name} and \kw{Area}, i.e., the co-authorship with each area for each researcher. We further slice-and-dice the results on \kw{Name} to only show the results related to ``Jiawei Han". We can find that in recent 20 years, Jiawei cooperated with researchers in \kw{DB} most. Meanwhile, in recent 10 years he has more and more cooperations with researchers in \kw{AI}, which can be ignored compared to cooperations with researchers in \kw{DB} in early 10 years.

\subsection{Efficiency Evaluation} \label{subsec:efficiency_eva}
In this section we mainly evaluate the efficiency of temporal graph cube on \kw{IMDB}, especially the efficiency of indexes.

\stitle{Exp-1: Average acceleration ratio of indexes with different similarity of snapshots.} The most important thing is how much indexes can accelerate the merging operation of snapshots in time ranges while conducting temporal cuboid queries. As we analysed before, the average aceleration ratio of indexes can vary if similarity of snapshots in temp-multi-networks changes. In this experiment, we only use \kw{IMDB} dataset, since vertices in \kw{IMDB} dataset have 7 dimensions and $ 2^7 $ views can be obtained. For convenience, we only materialize views in Table~\ref{tab:materialized_view},
\begin{table}[t!]%\vspace*{-0.3cm}
	%\scriptsize
	\small
	\centering
	\caption{Materialized views in experiments} \vspace*{-0.2cm} \label{tab:materialized_view}
	\begin{tabular}{c c c}
		\toprule
		\bf No. & Materialized Views & Similarity of Snapshots \cr %\hline %\hline
		\midrule
		1 & \kw{Pro} & 53.48 \cr %\hline
		\midrule
		2 & \kw{Pro}, \kw{Field} & 18.16 \cr %\hline
		\midrule
		3 & \kw{Pro}, \kw{Field}, \kw{Genre} & 6.15 \cr %\hline
		\midrule
		4 & \makecell[c]{\kw{Pro}, \kw{Field}, \kw{Genre},\\\kw{isAdult}}  & 6.07 \cr %\hline
		\midrule
		5 & \makecell[c]{\kw{Pro}, \kw{Field}, \kw{Genre},\\\kw{isAdult}, \kw{Averating}} & 4.70 \cr %\hline
		\midrule
		6 & \makecell[c]{\kw{Pro}, \kw{Field}, \kw{Genre},\\\kw{isAdult}, \kw{Averating}, \kw{Hotness}} & 3.97 \cr %\hline
		\midrule
		7 & \makecell[c]{\kw{Name}, \kw{Pro}, \kw{Field}, \kw{Genre},\\\kw{isAdult}, \kw{Averating}, \kw{Hotness}} & 1.12 \cr %\hline
		\bottomrule
	\end{tabular}
	\vspace*{-0.4cm}
\end{table}
among which we only materialize one view containing dimension \kw{Name} because in \kw{IMDB} dataset, \kw{Names} of two persons are hardly the same, so \kw{Name} can be approximately regarded as a primary key and having only one view containing dimension \kw{Name} is enough.

We will conduct a series of temporal cuboid queries on materialized views to examine the efficiency of indexes (conducting temporal cuboid queries on materialized views are all about merging snapshots in time ranges as we analyzed in \stitle{Temporal cuboid} part of Section~\ref{subsec:queries}). We manually set a series of time ranges to be specified in the queries, which are shown in Table~\ref{tab:time_ranges_set}.
\begin{table}[t!]%\vspace*{-0.3cm}
	%\scriptsize
	\small
	\centering
	\caption{Time ranges of queries in experiments} \vspace*{-0.2cm} \label{tab:time_ranges_set}
	\begin{tabular}{c c}
		\toprule
		\bf Length & Time ranges \cr %\hline %\hline
		\midrule
		4 & \makecell[c]{[1878,1880],[1891,1895],[1906,1910],\\{[1921,1925]},[1936,1940],[1951,1955],\\{[1966,1970]},[1981,1985],[1996,2000],[2011,2015]} \cr %\hline
		\midrule
		10 & \makecell[c]{[1878,1883],[1888,1898],[1903,1913],\\{[1918,1928]},[1933,1943],[1948,1958],\\{[1963,1973]},[1978,1988],[1993,2003],[2008,2018]} \cr %\hline
		\midrule
		20 & \makecell[c]{[1878,1888],[1883,1903],[1898,1918],\\{[1913,1933]},[1928,1948],[1943,1963],\\{[1958,1978]},[1973,1993],[1988,2008],[2003,2023]} \cr %\hline
		\midrule
		50 & \makecell[c]{[1878,1903],[1878,1918],[1883,1933],\\{[1898,1948]},[1913,1963],[1928,1978],\\{[1943,1993]},[1958,2008],[1973,2023],[1988,2028]} \cr %\hline
		\midrule
		100 & \makecell[c]{[1878,1928],[1878,1943],[1878,1958],\\{[1878,1973]},[1888,1988],[1903,2003],\\{[1918,2018]},[1933,2028],[1948,2028],[1963,2028]} \cr %\hline
		\bottomrule
	\end{tabular}
	\vspace*{-0.4cm}
\end{table}
The earliest and the latest timestamps in \imdb are 1878 and 2028 respectively. We set 5 groups of time ranges of different length: 4, 10, 20, 50, 100. For each length, we set 10 ranges which are evenly distributed in $ [1878,2028] $ as shown in Table~\ref{tab:time_ranges_set}. Ranges at two ends in each length group may be truncated since we try to keep each time range in $ [1878,2028] $. Specifically we set a series of queries $ QS_1,QS_2,...,QS_7 $, where $ QS_i $ is a set of queries specifying the No.$ i $ materialized view in Table~\ref{tab:materialized_view} and covering all time ranges of all lengths in Table~\ref{tab:time_ranges_set}, i.e., $ QS_1=\{((Pro),[1878,1880]),...,((Pro),[2011,2015]),\\((Pro),[1878,1883]),...,((Pro),[2008,2018]),...,\\((Pro),[1878,1928]),...,((Pro),[1963,2028])\} $.

We first set $ f_e= $ SUM($\ast$), and test the efficiency of prefix array (results are similar if $ f_e= $ COUNT($\ast$)). The results are shown in Table~\ref{tab:effi_prearray}.
\begin{table}[t!]%\vspace*{-0.3cm}
	%\scriptsize
	\small
	\centering
	\caption{Performance and acceleration ratio of prefix array on materialized views} \vspace*{-0.2cm} \label{tab:effi_prearray}
	\begin{tabular}{c c c c c}
		\toprule
		\makecell[c]{Views\\No.}& $ s $ & \makecell[c]{baseline\\(ms)} &\makecell[c]{pre-array\\(ms)}&\makecell[c]{acce-ratio\\(ave)} \cr %\hline %\hline
		\midrule
		1&53.48&17,272&3,216&4.30 \cr
		\midrule
		2&18.16&176,853&55,216&2.45\cr
		\midrule
		3&6.15&285,748&162,029&1.46\cr
		\midrule
		4&6.07&290,319&166,696&1.50\cr
		\midrule
		5&4.70&526,306&336,427&1.35\cr
		\midrule
		6&3.97&922,681&637,268&1.27\cr
		\midrule
		7&1.12&7,063,798&8,271,257&0.81\cr
		\bottomrule
	\end{tabular}
	\vspace*{-0.4cm}
\end{table}
%\begin{figure}[t] %\vspace*{-0.5cm}
%\begin{center}
%	\includegraphics[width=0.8\columnwidth]{fig/prearray_time.png}
%	\vspace*{-0.6cm}
%	\caption{prearray}
%	\vspace*{-0.4cm}
%	\label{fig:prearray_time}
%\end{center}
%\end{figure}
\kw{Views No.} is the No. of materialized views in Table~\ref{tab:materialized_view}, $ s $ is the similarity of snapshots. The third and the fourth columns are time used to conduct all prepared queries on each materialized view using baseline method and prefix array respectively, e.g., in the first line, conducting all queries in $ QS_1 $ takes $ 17,272ms $ using baseline method but $ 3,216ms $ using prefix array. \kw{acce-ratio(ave)} is the average acceleration ratio of prefix array on each query. In most cases, prefix array can accelerate merging snapshots in time ranges in conducting queries. However, the effect of acceleration varies with $ s $. $ s $ in the No.1 view is the biggest among the 7 views, and the average acceleration ratio of prefix array is also the biggest, while $ s $ in the No.7 view and the average acceleration ratio of prefix array in the No.7 view are both the lowest. We can easily find that the average acceleration ratio of prefix array drops if $ s $ drops in general, which confirms our analysis in Section~\ref{sec:effec_of_indexes}.

However, the similarity of snapshots is not the only difference among the 7 materialized views. In Table~\ref{tab:space_consumption_indexes} we can see that the size of materialized views is also different, so in our experiment, we choose the No.5 view as a reference and manually construct some temp-multi-networks, which have different similarity of snapshots but the same length of snapshot array and number of edges in total with the No.5 view. The numbers of edges in snapshots of each man-made dataset are almost uniformly distributed. The results on man-made datasets are shown in Table~\ref{tab:effi_prearray_man_made}.
\begin{table}[t!]%\vspace*{-0.3cm}
	%\scriptsize
	\small
	\centering
	\caption{Performance and acceleration ratio of prefix array on man-made datasets} \vspace*{-0.2cm} \label{tab:effi_prearray_man_made}
	\begin{tabular}{c c c c c}
		\toprule
		\makecell[c]{Original\\size}& $ s $ & \makecell[c]{baseline\\(ms)} &\makecell[c]{pre-array\\(ms)}&\makecell[c]{acce-ratio\\(ave)} \cr %\hline %\hline
		\midrule
		5,467,826 &10&589,636 &442,730 &1.20 \cr
		\midrule
		5,467,826 &20&448,802 &234,486 &1.80 \cr
		\midrule
		5,467,826 &50&327,947 &98,095 &3.35 \cr
		\midrule
		5,467,826 &80&306,467 &64,689 &4.80 \cr
		\bottomrule
	\end{tabular}
	\vspace*{-0.4cm}
\end{table}
%\begin{figure}[t] %\vspace*{-0.5cm}
%	\begin{center}
%		\includegraphics[width=0.8\columnwidth]{fig/create_prearray_time.png}
%		\vspace*{-0.6cm}
%		\caption{man-made prearray time}
%		\vspace*{-0.4cm}
%		\label{fig:man_made_prearray_time}
%	\end{center}
%\end{figure}
Each data line in Table~\ref{tab:effi_prearray_man_made} represents result on each man-made temp-multi-network. For example, in the first data line, the man-made temp-multi-network has 5,467,826 edges as the No.5 materialized view but has similarity of snapshots of 10. To conduct similar queries in $ QS_1,...,QS_7 $, baseline method uses 589,636 ms and prefix array uses 442,730 ms, obtaining an acceleration ratio of 1.2. In the last data line, the man-made temp-multi-network has the same number of edges but larger similarity of snapshots of 80. Prefix array in this case obtains an acceleration ratio of 4.8. We can see that the results here are consistent with results in Table~\ref{tab:effi_prearray}. Note that the difference of time consumption using baseline method among man-made temp-multi-networks may be due to the implementation of the program and it doesn't affect our results.

We then set $ f_e= $ MAX($\ast$) and use the same queries in $ QS_1,...,QS_7 $ above to test the efficiency of STable and STree (it's similar if $ f_e= $ MIN($\ast$)). The results are shown in Table~\ref{tab:effi_stree_stable}.
\begin{table*}[t!]%\vspace*{-0.3cm}
	%\scriptsize
	\small
	\centering
	\caption{Performance and acceleration ratio of STree and STable on materialized views} \vspace*{-0.2cm} \label{tab:effi_stree_stable}
	\begin{tabular}{c c c c c c c}
		\toprule
		\makecell[c]{Views\\No.}& $ s $ & \makecell[c]{baseline\\(ms)} &\makecell[c]{STable\\(ms)}&\makecell[c]{acce-ratio\\(STable, ave)}&\makecell[c]{STree\\(ms)}&\makecell[c]{acce-ratio\\(STree, ave)} \cr %\hline %\hline
		\midrule
		1&53.48&16,960 &1,313 &10.75 &3,960 &3.45 \cr
		\midrule
		2&18.16&172,614 &21,921 &6.30 &57,637 &2.32 \cr
		\midrule
		3&6.15&279,684 &54,267 &4.48 &144,849 &1.59 \cr
		\midrule
		4&6.07&285,623 &55,414 &4.00 &147,565 &1.59 \cr
		\midrule
		5&4.70&515,607 &106,463 &3.82 &293,546 &1.50 \cr
		\midrule
		6&3.97&896,386 &195,698 &3.62 &550,376 &1.40 \cr
		\midrule
		7&1.12&6,719,417 &NaN&NaN&6,366,660 &1.05 \cr
		\bottomrule
	\end{tabular}
	\vspace*{-0.4cm}
\end{table*}
%\begin{figure}[t] %\vspace*{-0.5cm}
%	\begin{center}
%		\includegraphics[width=0.8\columnwidth]{fig/stree_and_stable_time.png}
%		\vspace*{-0.6cm}
%		\caption{stree and stable}
%		\vspace*{-0.4cm}
%		\label{fig:stree_and_stable_time}
%	\end{center}
%\end{figure}
We can see that both STree and STable can accelerate the queries, while the effect of STable is more significant, since there are at most 2 snapshots being operated in STable for every query. However, we failed to build STable on the No.7 view because of the huge memory consumption, which we will discuss in Exp-2. Similar to prefix array, the effects of STree and STable are influenced by the similarity of snapshots, e.g., on the No.1 view, STree is $ 3.45 $ times faster than baseline method and STable is $ 12.92 $ times faster than baseline method. But when similarity of snapshots in views drops, both STree and STable gain poorer results, e.g., on the No.7 view, STree is almost unhelpful, and STable can not even be built.

To eliminate the influence of different size of materialized views, we build indexes on our man-made temp-multi-networks and conduct similar queries in $ QS_1,...,QS_7 $. The results are shown in Table~\ref{tab:effi_stree_stable_man_made},
\begin{table*}[t!]%\vspace*{-0.3cm}
	%\scriptsize
	\small
	\centering
	\caption{Performance and acceleration ratio of STree and STable on man-made datasets} \vspace*{-0.2cm} \label{tab:effi_stree_stable_man_made}
	\begin{tabular}{c c c c c c c}
		\toprule
		\makecell[c]{Original\\size}& $ s $ & \makecell[c]{baseline\\(ms)} &\makecell[c]{STable\\(ms)}&\makecell[c]{acce-ratio\\(STable, ave)}&\makecell[c]{STree\\(ms)}&\makecell[c]{acce-ratio\\(STree, ave)} \cr %\hline %\hline
		\midrule
		5,467,826 &10&359,416 &127,685 &2.77 &280,001 &1.26 \cr
		\midrule
		5,467,826 &20&298,525 &81,443 &3.83 &197,733 &1.58 \cr
		\midrule
		5,467,826 &50&239,775 &36,002 &6.66 &96,798 &2.43 \cr
		\midrule
		5,467,826 &80&225,354 &23,234 &9.56 &67,257 &3.20 \cr
		\bottomrule
	\end{tabular}
	\vspace*{-0.4cm}
\end{table*}
%\begin{figure}[t] %\vspace*{-0.5cm}
%	\begin{center}
%		\includegraphics[width=0.8\columnwidth]{fig/create_stree_and_stable_time.png}
%		\vspace*{-0.6cm}
%		\caption{man-made stree and stable time}
%		\vspace*{-0.4cm}
%		\label{fig:man_made_stree_and_stable_time}
%	\end{center}
%\end{figure}
which are similar to results in Table~\ref{tab:effi_stree_stable}. The average acceleration ratio of STree and STable both increases when similarity of snapshots increases, and the effect of STable is still more significant.

\stitle{Exp-2: Space consumption of indexes with different similarity of snapshots.} In this experiment we mainly discuss the space consumption of indexes.
\begin{table*}[t!]%\vspace*{-0.3cm}
	%\scriptsize
	\small
	\centering
	\caption{Space consumption of indexes on materialized views} \vspace*{-0.2cm} \label{tab:space_consumption_indexes}
	\begin{tabular}{c c c c c c c c c}
		\toprule
		\makecell[c]{Views\\No.}&\makecell[c]{Original\\size}& $ s $ &\makecell[c]{pre-array\\size}&\makecell[c]{pre-array\\ratio}& \makecell[c]{STable\\size} &\makecell[c]{STable\\ratio}&\makecell[c]{STree\\size}&\makecell[c]{STree\\ratio} \cr %\hline %\hline
		\midrule
		1&41,448 &53.48&62,535 &1.51 &388,772 &9.38&93,975 &2.27   \cr
		\midrule
		2&396,755 &18.16&1,022,351 &2.58  &5,734,822 &14.45&1,077,456 &2.72  \cr
		\midrule
		3&3,738,125 &6.15&21,518,053 &5.76  &100,111,214 &26.78&13,788,570 &3.69  \cr
		\midrule
		4&3,758,885 &6.07&21,949,182 &5.84  &101,640,912 &27.04&13,918,066 &3.70  \cr
		\midrule
		5&5,467,826 &4.70&39,066,700 &7.14  &173,573,474 &31.74&22,063,352 &4.04  \cr
		\midrule
		6&8,304,421 &3.97&69,167,227 &8.33  &297,747,375 &35.85&35,837,987 &4.32  \cr
		\midrule
		7&37,096,632 &1.12&1,015,573,783 &27.38  &NaN&NaN&285,623,583 &7.70 \cr
		\bottomrule
	\end{tabular}
	\vspace*{-0.4cm}
\end{table*}
%\begin{figure}[t] %\vspace*{-0.5cm}
%	\begin{center}
%		\includegraphics[width=0.8\columnwidth]{fig/index_space.png}
%		\vspace*{-0.6cm}
%		\caption{indexes size}
%		\vspace*{-0.4cm}
%		\label{fig:index_size}
%	\end{center}
%\end{figure}
Table~\ref{tab:space_consumption_indexes} shows the space consumption of each index on all materialized views. \kw{Original} \kw{size} is the total number of edges in materialized views. \kw{pre}\kw{-}\kw{array} \kw{size} is the total number of edges in prefix array built on each view. \kw{pre}\kw{-}\kw{array} \kw{ratio} is the space consumption ratio of prefix array built on each view. \kw{STable} \kw{size}, \kw{STable} \kw{ratio}, \kw{STree} \kw{size} and \kw{STree} \kw{ratio} are similar. We can see that the space consumption ratio of each index grows when similarity of snapshots drops, which confirms our conjecture in Section~\ref{sec:effec_of_indexes}. In detail, STree is the most stable and cost-optimal index among the three indexes, because the space consumption ratio of STree is relatively low and has the slowest growth when similarity of snapshots drops. Combining the results in Exp-1, STree achieves a good balance between time and space. STable here has the highest space consumption ratio, since it maintains two-dimensional snapshot array. STable can not be built on the No.7 view since it takes a lot of time and space.

Similarly, we conduct the same experiment on our man-made datasets to explore the relationship between space consumption ratio of indexes and similarity of snapshots again. The results are shown in Table~\ref{tab:space_consumption_indexes_man_made},
\begin{table*}[t!]%\vspace*{-0.3cm}
	%\scriptsize
	\small
	\centering
	\caption{Space consumption of indexes on man-made datasets} \vspace*{-0.2cm} \label{tab:space_consumption_indexes_man_made}
	\begin{tabular}{c c c c c c c c}
		\toprule
		\makecell[c]{Original\\size}& $ s $ &\makecell[c]{pre-array\\size}&\makecell[c]{pre-array\\ratio}& \makecell[c]{STable\\size} &\makecell[c]{STable\\ratio}&\makecell[c]{STree\\size}&\makecell[c]{STree\\ratio} \cr %\hline %\hline
		\midrule
		5,467,826 &10&73,934,678 &13.52 &238,281,300 &43.58 &28,776,691 &5.26 \cr
		\midrule
		5,467,826 &20&40,179,416 &7.35 &169,629,213 &31.02 &25,472,457 &4.66 \cr
		\midrule
		5,467,826 &50&16,362,842 &2.99 &86,337,728 &15.79 &18,116,929 &3.31 \cr
		\midrule
		5,467,826 &80&10,276,491 &1.88 &59,063,641 &10.80 &14,966,701 &2.74 \cr
		\bottomrule
	\end{tabular}
	\vspace*{-0.2cm}
\end{table*}
%\begin{figure}[t] %\vspace*{-0.5cm}
%	\begin{center}
%		\includegraphics[width=0.8\columnwidth]{fig/create_index_size.png}
%		\vspace*{-0.6cm}
%		\caption{create indexes size}
%		\vspace*{-0.4cm}
%		\label{fig:create_index_size}
%	\end{center}
%\end{figure}
which are similar to results in Table~\ref{tab:space_consumption_indexes}, so we don't describe them again.

\stitle{Exp-3: Updating performance of indexes with different similarity of snapshots.} In this experiment, we mainly discuss the updating performance of each index built on views with different similarity of snapshots, specifically the trend of time and space consumption under continuous addition of new edges. However, our datasets (DBLP, IMDB and man-made datasets) have no more edges which can be added after we build temporal graph cubes on them. In this experiment we simulate the continuous addition of new edges on indexes by dividing all temporal edges in a materialized view into 10 batches equally according to the order of timestamps, e.g., the first batch of temporal edges are edges with the smallest timestamps, and they are added into indexes one by one according to the order of timestamps. When edges in a single batch are added into indexes, we record the total time consumption of adding edges in the batch and current size, represented by the number of stored edges in the indexes. As the previous experiments, we use man-made datasets to eliminate the difference on size of materialized views. Due to the limitation of space, we directly conduct experiments on man-made datasets.

Similar to previous experiments, we construct 4 man-made datasets having the same number of temporal edges with the No.5 view but different similarity of snapshots, which are denoted by $ s=10 $, $ s=20 $, $ s=50 $ and $ s=80 $.
\begin{figure*}[t!]
	\begin{center}
		\begin{tabular}[t]{c}\hspace*{-0.3cm}
			\subfigure[{\scriptsize Prefix Array}]{
				\includegraphics[width=0.25\textwidth]{fig/exp/index_update/index_update_single_a.pdf}
			}
			\hspace*{-0.3cm}
			\subfigure[{\scriptsize STable}]{
				\includegraphics[width=0.25\textwidth]{fig/exp/index_update/index_update_single_b.pdf}
			}
			\hspace*{-0.3cm}
			%\vspace*{-0.3cm}\\\hspace*{-0.3cm}
			\subfigure[{\scriptsize STree}]{
				\includegraphics[width=0.25\textwidth]{fig/exp/index_update/index_update_single_c.pdf}
			}
			\\
			\hspace*{-0.3cm}
			\subfigure[{\scriptsize Prefix Array}]{
				\includegraphics[width=0.25\textwidth]{fig/exp/index_update/index_update_single_d.pdf}
			}
			\hspace*{-0.3cm}
			\subfigure[{\scriptsize STable}]{
				\includegraphics[width=0.25\textwidth]{fig/exp/index_update/index_update_single_e.pdf}
			}
			\hspace*{-0.3cm}
			\subfigure[{\scriptsize STree}]{
				\includegraphics[width=0.25\textwidth]{fig/exp/index_update/index_update_single_f.pdf}
			}
		\end{tabular}
	\end{center}
	\vspace*{-0.2cm}
	\caption{Trend of time and space consumption of indexes under continuous addition of new edges}
	\vspace*{-0.4cm}
	\label{fig:index_update}
\end{figure*}

In Figure~\ref{fig:index_update}, (a)-(c) show the trend of time consumption of each index being added with batches of edges in the 4 man-made datasets.
%We transfer the actual value of time consumption to the ratio of time consumption value of adding each batch of edges to time consumption value of adding the 1-st batch of edges, for example, on curve related to $ s=10 $ in Figure~\ref{fig:index_update} (a), time consumption value of adding the 3-th batch of edges to prefix array is about 1.6 times of time consumption value of adding the 1-st batch to the same prefix array.
Overall, in Figure~\ref{fig:index_update} (b)-(c), we can see that all four curves in these two figures show an increase, which means the time consumption of adding the back batch of edges of each dataset into both STable and STree are more than adding the front batch of edges. This pattern can be explained by our analysis in Section~\ref{subsec:sparse_table} and \ref{subsec:segment_tree}. However, after a rise, some curves in Figure~\ref{fig:index_update} (a) show a small decline and eventually stabilize, while curves related to $ s=50 $ and $ s=80 $ do not even have a rising process. It can still be explained that, in the final stabilization phase of curves, the last snapshot of the prefix array already shares the same edges with the summative snapshot of snapshot array in temp-multi-network the prefix array built on. In this phase, the last snapshot of prefix array to be copied ($ preArray[t_{end}] $ in line~2 of Algorithm~\ref{alg:prefix_array_update}) has a constant size when adding new edges with $ t-t_{base}>t_{end} $, so the total time consumption per batch is constant. As for the small decline of curves, it's because in the stabilization phase, regardless of whether $ preArray[t_{end}] $ is copied beforehand, the new edge already exists in the last snapshot of the prefix array when it's finally added, so the branch of Algorithm~\ref{alg:prefix_array_update} in line~5-6 will not be executed, which creates edge and takes more time than branch in line~3-4 in actual running. It's easy to know that the larger $ s $ is, the earlier the last snapshot of the prefix array shares the same edges with the summative snapshot, which is shown in Figure~\ref{fig:index_update} (a).

Apart from the overall trend of the curves, in Figure~\ref{fig:index_update} (a)-(c), we can find that in each figure, higher curves are related to smaller values of $ s $, which means for each type of index, when adding the same number of new edges, indexes built on datasets with smaller similarity of snapshots take more time to be updated. For example in Figure~\ref{fig:index_update} (a), when adding the same batch of edges, the last snapshot of the prefix array will be copied for the new edges with $ t-t_{base}>t_{end} $. When $ s $ is small, the last snapshot of the prefix array built on such a dataset has larger size because snapshots in snapshot array of the original dataset share fewer edges. In this case, updating prefix array takes more time. Similar reason can be used to explain Figure~\ref{fig:index_update} (b). Updating STree also involves copying snapshot, i.e., snapshot in root, which maintains all snapshots in snapshot array, and gets larger if $ s $ gets smaller. However, copying snapshot occurs only when creating new root for STree, and since the range maintained by the root grows exponentially, copying snapshot hardly occurs when adding new edges in later batches. The difference in time consumption when adding the same batch of new edges of datasets with different similarity of snapshots $ s $ mainly comes from updating the root down to the leaf with new edges. It can be speculated from probability that for STree built on datasets with larger $ s $, there is a higher possibility that the snapshot in each node being updated with the new edge already contains that new edge, so the time consumption is slightly less. The deeper reason is that more edges are shared among snapshots of the original dataset if $ s $ of the dataset is larger. The time consumption of creating these nodes is the same if they do not yet exist. Since copying snapshot rarely occurs, the time consumption of updating STree is least influenced by $ s $ of the datasets among the three indexes, which can be shown in Figure~\ref{fig:index_update} (c).

Figure~\ref{fig:index_update} (d)-(f) show the trend of space consumption of each index being added with each batch of edges of the 4 man-made datasets. Overall, curves in Figure~\ref{fig:index_update} (d) and (f) are linear, while curves in Figure~\ref{fig:index_update} (e) grow faster than linear curve, because STable maintains a two-dimensional snapshot array, which we discussed in Section~\ref{subsec:sparse_table}. The linear curves in Figure~\ref{fig:index_update} (d) and (f) show that prefix array and STree have better scalability than STable. For each kind of index, the instances built on datasets with smaller $ s $ have more space consumption and grow faster after each batch of new edges are added. The reason is also related to $ s $. For example in Figure~\ref{fig:index_update} (d), when adding each batch of new edges, prefix array built on datasets with smaller $ s $ has to copy larger snapshot. Among these three indexes, the space consumption of updating STree is still least influenced by $ s $ of the datasets because of the rare occurrence of copying snapshot and close space consumption caused by other operations in updating STree built on datasets with different $ s $. The relatively flat curves in Figure~\ref{fig:index_update} (f) also indicate that STree has the best scalability.

\section{Related Work} \label{sec:related_work}
Our work extends data warehouse and OLAP techniques to temporal multidimensional networks. There are many works applying data warehouse and OLAP to other data types, such as stream data \cite{han2005stream}, spatio-temporal data \cite{gomez2009survey}, sequence data \cite{lo2008olap,liu2011cube}, and text data \cite{lin2008text}. Except the traditional relational data, Dehdouh et al. \cite{dehdouh2014columnar} tried to compute data cubes from column-oriented NoSQL data. As for graph data, except \cite{zhao2011graph}, some other works focus on heterogeneous networks \cite{yin2012hmgraph,wang2015tsmh}, but none of them focus on temporal multidimensional networks.

Our work is also related to graph summarization and analysis. For static graph, there are some representative works, including graph compression with MDL (Minimum Description Length) principle \cite{navlakha2008graph}, attributed graph compression \cite{wu2014graph} and graph clustering based on vertices partitioning \cite{zhou2009graph}. In recent years there are also some novel works related to static graph summarization, such as summarizing directed acyclic graph (DAG) \cite{zhu2020top}, summarizing multi-relation graph \cite{ke2022multi} where multiple edges of different types may exist between any pair of vertices, and DPGS model \cite{zhou2021dpgs}, which can preserve properties like graph spectrum and the authorities and hubnesses of vertices especially in highly degree-skewed graphs while reconstructing graph. For temporal graph and dynamic graph, TimeCrunch \cite{shah2015timecrunch} summarizes a large dynamic graph with a set of important temporal structures using MDL, such as ranged full clique, periodic bipartite core, oneshot star, etc. \cite{fernandes2018dynamic,tsalouchidou2018scalable} both investigate the problem of summarizing dynamic graph in real time. It is done by using a sliding window of fixed length to summarize the latest snapshots. When a new snapshot is generated, the window slides forward, discarding the oldest snapshot in the window and generating a new summary. However, in this paper, we can specify arbitrary time window to query summative result in real time. In addition, we summarize the graph based on the selected dimensions of vertices, while the previous two works summarize the graph based on the super-node and super-edge. MoSSo \cite{ko2020incremental} incrementally summarizes fully dynamic graph. The difference of \cite{ko2020incremental} with the previous two works is that the edges in fully dynamic graph are dynamically added and removed, and MoSSo incrementally summarize the whole graph after any modification on the graph.

\section{Conclusion} \label{sec:conclusion}
In this paper, we extend data warehouse and OLAP technology to temporal multidimensional networks by proposing a new data warehouse model \kw{Temporal\enspace Graph\enspace Cube}, which provides knowledge workers with tools to analyze temporal multidimensional networks. We first extend the basic concepts in static graph cube and discuss two basic queries, \kw{temporal\enspace cuboid} and \kw{temporal\enspace crossboid}, allowing \kw{Temporal\enspace Graph\enspace Cube} to support OLAP queries on temporal multidimensional networks. Then we extend three typical indexes on common range query problem to accelerate the above queries. We find a simple metric that affects the efficiency of indexes and analyze how it affects. We point out that some classical methods are no longer applicable in the implementation of \kw{Temporal\enspace Graph\enspace Cube}. Our experimental results show the effectiveness of \kw{Temporal\enspace Graph\enspace Cube} and corroborate our inferences about the relationship between the efficiency of indexes and the metric. With the support of the implementation strategy for \kw{Temporal\enspace Graph\enspace Cube}, the indexes are able to achieve an acceleration effect. Moreover, the best time and space balance, as well as the best scalability, is achieved by segment tree among the three indexes.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

% use section* for acknowledgment
\ifCLASSOPTIONcompsoc
% The Computer Society usually uses the plural form
\section*{Acknowledgments}
\else
% regular IEEE prefers the singular form
\section*{Acknowledgment}
\fi
This work was partially supported by (i) National Key Research and Development Program of China 2020AAA0108503, (ii) NSFC Grants 62072034 and 61772346, and (iii) CCF-Baidu Open Fund;
%(ii) NSFC Grants 61772346, 61732003, U1809206, 61772091, 61802035;
%(iii) Beijing Institute of Technology Research Fund Program for Young Scholars;
%(iv) Research Grants Council of the Hong Kong SAR, China No. 14221716 and 14203618;
%(v) ARC Discovery Project Grant DP160101513;
%and (vi) MOE, Singapore under grant MOE2015-T2-2-069 and NUS, Singapore under an SUG.

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
\newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)

%\balance
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,temporal_graph_cube}

%\begin{thebibliography}{1}
%
%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
%
%\end{thebibliography}

% biography section
%
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:
\vspace*{-20mm}

% if you will not have a photo at all:
%\begin{IEEEbiographynophoto}{Rong-Hua Li}
%Biography text here.
%\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

%\begin{IEEEbiographynophoto}{Jane Doe}
%Biography text here.
%\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}